---
title: "The penguins of Antarctica"
author: "Selina Baldauf"
date: "25/3/2022"
output: 
  pdf_document: 
    number_sections: true
    toc: true
    fig_caption: true
    df_print: "kable"
  word_document:
    number_sections: true
    toc: true
    fig_caption: true
    df_print: "kable"
  html_document:
    number_sections: true
    toc: true
    toc_float: true
    code_folding: "hide"
    code_download: true
    highlight: "textmate"
    theme: "journal"
    df_print: "kable"
bibliography: references.bib
urlcolor: "blue"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, # code is included in output
  warning = FALSE, # don't show warnings
  message = FALSE # don't show messages
)
library(tidyverse)
library(palmerpenguins)
```

# Introduction

There are three main penguin species in Antarctica (*Chinstrap*, *Gentoo*, *Adelie*). You can see them in the following figure:

<!-- ![Artwork by [Allison Horst](https://twitter.com/allison_horst)](https://github.com/allisonhorst/palmerpenguins/blob/master/man/figures/lter_penguins.png?raw=true) -->

In this paper we want to answer the following questions

1.  How bill depth depends on bill length?
2.  Which penguin species has the highest body mass?

# Methods

All analysis was done using R version 4.1.3 [@Rcore2022] and the R markdown package [@Allaire2021].

## The data

The data was collected on islands in Antarctica and published by @gorman2014. You can find the original paper with the title "Ecological sexual dimorphism and environmental variability within a community of Antarctic penguins (genus *Pygoscelis*)" [@gorman2014] in PLoS ONE[^1]

[^1]: paper available [here](https://doi.org/10.1371/journal.pone.0090081).

The data is published via the `palmerpenguins` R package [@horst2020] which you can find [on this website](https://allisonhorst.github.io/palmerpenguins/).

**The data contains (among others) the following measurements:**

-   bill length

-   bill depth

-   body mass

-   sex

    -   male
    -   female

## The analysis

We did some plots, calculated some summary statistics and a linear model of the form $y = ax + b + \epsilon$

# Results

```{r mean-penguin-species, include=FALSE}
# calculate the mean body mass of the penguin species
adelie_mean <- penguins %>% filter(species == "Adelie") %>% pull(body_mass_g) %>% mean(na.rm = TRUE) 
chinstrap_mean <- penguins %>% filter(species == "Chinstrap") %>% pull(body_mass_g) %>% mean(na.rm = TRUE) 
gentoo_mean <- penguins %>% filter(species == "Gentoo") %>% pull(body_mass_g) %>% mean(na.rm = TRUE) 
```

The mean weight of all penguin species is `r mean(penguins$body_mass_g, na.rm = TRUE)`. *Gentoo* penguins have an average weight of `r round(gentoo_mean,0)` g, *Adelie* penguins of `r round(adelie_mean, 0)` g and *Chinstrap* penguins of `r round(chinstrap_mean, 0)` g.

The figure below shows that *Gentoo* penguins have the highest body mass.

```{r histogram-weight, fig.cap="Histogram of weight of the three penguin species.", fig.width=4, fig.height=3, fig.align='center'}
ggplot(penguins, aes(x = body_mass_g, fill = species)) +
  geom_histogram(alpha = 0.6) +
   scale_fill_manual(values = c("darkorange", "purple", "cyan4")) +
  theme_minimal()
```

There is a positive relationship between bill length and bill depth for all 3 species, as the figure below shows.

```{r scatterplot-bill, eval=FALSE, fig.cap="Scatter plot with regression lines showing the relationship between bill length and bill depth for the 3 penguin species"}
ggplot(
  data = penguins,
  aes(
    x = bill_length_mm,
    y = bill_depth_mm,
    color = species,
    shape = species
  )
) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  theme_minimal()
```

In general, it looks like the body characteristics differ between the sexes but also between the penguin species, as the table below illustrates:

```{r summary-table}
penguins %>% 
  filter(!(is.na(sex))) %>% 
  group_by(species, sex) %>% 
  summarize(
    bill_length = mean(bill_length_mm, na.rm = TRUE),
    bill_depth = mean(bill_depth_mm, na.rm = TRUE),
    flipper_length = mean(flipper_length_mm, na.rm = TRUE),
    body_mass = mean(body_mass_g, na.rm = TRUE)
  )
```

# References
