---
title: "Task 2: Beautify and save plots"
number-sections: false
---

[← Back to session page](../07_ggplot.qmd)

:::{.callout-note}

## Find the solution here after the session ends.

```{=html}
<script>const course_day = 2;</script>
<script src="../../timer_solution.js"></script>
<p class="r_class_solution" style="display:none">
  <a href="../solutions/ggplot-solution-2.html" target="_blank">Click here for the solution.</a>
</p>
```
:::

## Beautify the plots

Pick one of the tasks below first, then move on to saving your plot. Feel free to try your own ideas too.

### Beautify a plot from Task 1

Take a plot from Task 1 and make it look nicer. Here are some ideas:

- Add a theme layer (e.g. `theme_minimal()`)
- Customize the theme, e.g. change the legend position or make axis titles bold
- Change the color/fill scale using `scale_color_manual` or `scale_color_viridis_d()`
- Add proper axis labels and a title with `labs()`
- Make points transparent, give them a different shape, ...

### Reproduce this plot

Try to reproduce the following plot. It builds on the scatterplot from Task 1:

```{r fig.cap="Visualization from the [palmerpenguins package](https://allisonhorst.github.io/palmerpenguins/)"}
#| echo: false
knitr::include_graphics(here::here(
  "sessions/slides/img/figs-penguin/README-flipper-bill-1.png"
))
```

Hint: the colors used are:

```{r echo=TRUE, eval=FALSE}
penguin_cols <- c("darkorange", "purple", "cyan4")
```

## Save your plot

Take one of your plots and save it to your project as a `.png` file.

1. Make sure you have an `img/` folder in your project (create it in the Files pane if needed)
2. Use `ggsave()` to save your plot as a `.png` file in the `img/` folder. Check `?ggsave` or the slides if you're not sure about the arguments.
3. Check the Files pane to find the png file.

## References

Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0.
https://allisonhorst.github.io/palmerpenguins/. doi: 10.5281/zenodo.3960218.
