---
title: "Task 2: Beautify and save plots"
number-sections: false
---

[← Back to session page](../07_ggplot.qmd)

:::{.callout-note}

## Find the solution here after the session ends.

```{=html}
<script>const course_day = 2;</script>
<script src="../../timer_solution.js"></script>
<p class="r_class_solution" style="display:none">
  <a href="../solutions/ggplot-solution-2.html" target="_blank">Click here for the solution.</a>
</p>
```
:::

## Getting started

Take one of the plots you made in [Task 1](ggplot-task-1.qmd) and improve it step by step.

First, save the plot in a variable so you can build on it without rewriting the whole plot each time:

```{r}
#| eval: false
#| echo: true
my_plot <- ggplot(...) + geom_...(...)
my_plot

# Try something new
my_plot + ...

# Keep the change
my_plot <- my_plot + ...
my_plot
```

:::{.callout-note collapse="true"}

## I don't have a working plot from Task 1 (unfold to get a starting plot)

Copy and paste this code to get a starting plot:

```{r}
#| eval: false
#| echo: true
library(tidyverse)

ggplot(
  penguins,
  aes(x = bill_len, y = bill_dep, color = species, shape = species)
) +
  geom_point()
```

:::

## Add labels

Use `labs()` to give your plot a title and proper axis labels. Also change the legend title so it doesn't just show the variable name.

## Change the colors

Change the colors of your plot. You can either:

- Pick your own colors with `scale_color_manual()`. Refer to colors by name (e.g. `"red"`, `"cyan4"`) or by hex code (e.g. `"#FF0000"`). You can find hex codes with this [color picker](https://www.w3schools.com/colors/colors_picker.asp) and all R color names [here](https://derekogle.com/NCGraphing/resources/colors).
- Use a built-in palette with `scale_color_brewer()` or `scale_color_viridis_d()`.

:::{.callout-tip}

## Color vs. Fill

If your plot uses `fill` (e.g. boxplots, histograms), you need `scale_fill_manual()` or `scale_fill_brewer()` instead of the `scale_color_*` versions.

:::

## Apply a theme

Try a few different themes (e.g. `theme_minimal()`, `theme_classic()`, `theme_bw()`) and pick one you like.

## Reproduce this plot

Try to reproduce the following plot. It builds on the scatterplot from Task 1:

```{r fig.cap="Visualization from the [palmerpenguins package](https://allisonhorst.github.io/palmerpenguins/)"}
#| echo: false
knitr::include_graphics(here::here(
  "sessions/slides/img/figs-penguin/README-flipper-bill-1.png"
))
```

Hint: the colors used are:

```{r echo=TRUE, eval=FALSE}
penguin_cols <- c("darkorange", "purple", "cyan4")
```

## Save your plot

Take one of your plots and save it to your project as a `.png` file.

1. Make sure you have an `img/` folder in your project (create it in the Files pane if needed)
2. Use `ggsave()` to save your plot as a `.png` file in the `img/` folder. Check `?ggsave` or the slides if you're not sure about the arguments.
3. Check the Files pane to find the png file.

## For the fast ones

You can do these in any order, or skip them and just take a break.

### Fine-tune the theme

Use `theme()` to customize individual elements of your plot, for example the legend position or the font style of the title. To find out what you can customize:

- Check `?theme`
- Search online "ggplot theme ..." for specific ideas
- Have a look at the theme reference sheet on the slides

### Change point appearance

Try setting fixed aesthetics inside `geom_point()` to change the look of all points, for example their size, transparency (`alpha`), or shape.

### Save in different formats and sizes

Try saving your plot as a `.pdf` and with custom dimensions using `width` and `height` in `ggsave()`.

### Make a plot with your own data

If you have your own data loaded in R, try making a plot with it and beautifying it.

## References

Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0.
https://allisonhorst.github.io/palmerpenguins/. doi: 10.5281/zenodo.3960218.
