---
title: "Task: Reproducible documents with Quarto"
number-sections: false
---

[← Back to session page](../11_quarto.qmd)

:::{.callout-note}

## Find the solution here after the session ends.

```{=html}
<script>const course_day = 3;</script>
<script src="../../timer_solution.js"></script>
<p class="r_class_solution" style="display:none">
  <a href="../solutions/quarto-solution.html" target="_blank">Click here for the solution.</a>
</p>
```
:::

## Create your first Quarto document

**1.** In the **Files pane**, click `New File -> Quarto Document`. Give it a name like `penguin_report.qmd`.

**2.** The file is empty. Add a YAML header at the top with a title, your name, and the output format:

```yaml
---
title: "Penguin Report"
author: "Your Name"
format: html
---
```

## Add text

**4.** Below the YAML header, add a section header and a short sentence:

```md
## Introduction

This is a short report about the **Palmer Penguins** dataset.
```

## Add code

**5.** Add a code chunk that loads the tidyverse package. Hide the messages by adding chunk options:

````md
```{{r}}
#| message: false
#| warning: false

library(tidyverse)
```
````

**6.** Add a code chunk that prints the first rows of the penguins data:

````md
```{{r}}
head(penguins)
```
````

**7.** Add a new section header `## Plot` and a code chunk that creates a scatterplot:

````md
```{{r}}
ggplot(penguins, aes(x = bill_len, y = bill_dep, color = species)) +
  geom_point() +
  labs(
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    color = "Species"
  ) +
  theme_minimal()
```
````

## Add inline code

**8.** Below the plot, add a sentence that uses inline code:

```md
The dataset contains `r knitr::inline_expr("nrow(penguins)")` penguins 
from `r knitr::inline_expr("length(unique(penguins$species))")` species.
```

## Render

**9.** Click the **Render** button (or press `Ctrl + Shift + K`) to render your document to HTML. Check the result!

## For the fast ones

You can do these in any order, or skip them and just take a break.

### Add a table of contents

Add `toc: true` to your YAML header under the html format and re-render:

```yaml
---
title: "Penguin Report"
author: "Your Name"
format:
  html:
    toc: true
---
```

### Add a figure caption

Add a caption to your plot chunk using the `#| fig-cap:` option. Re-render and check how it looks.

### Hide the code

Try adding `code-fold: true` under the html format in the YAML header. This lets readers click to expand the code.

### Add another section

Add a new section to your report. For example, a summary table or a second plot. Here is some code you could use:

A boxplot of body mass by species:

```{r}
#| eval: false
#| echo: true
ggplot(penguins, aes(x = species, y = body_mass, fill = species)) +
  geom_boxplot() +
  labs(x = "Species", y = "Body mass (g)") +
  theme_minimal() +
  theme(legend.position = "none")
```

A summary table:

```{r}
#| eval: false
#| echo: true
penguins |>
  group_by(species) |>
  summarize(
    n = n(),
    mean_mass = mean(body_mass, na.rm = TRUE),
    mean_flipper = mean(flipper_len, na.rm = TRUE)
  )
```

### Render to PDF

Change `format: html` to `format: pdf` and render. If you get an error about missing LaTeX, run this in the R console first:

```{r}
#| eval: false
#| echo: true
quarto::quarto_install_tinytex()
```
