---
title: "Task: Import and Export Data"
number-sections: false
code-line-numbers: false
---

[← Back to session page](../05_readr.qmd)

:::{.callout-note}

## Find the solution here after the session ends.

```{=html}
<script>const course_day = 1;</script>
<script src="../../timer_solution.js"></script>
<p class="r_class_solution" style="display:none">
  <a href="../solutions/readr.html" target="_blank">Click here for the solution.</a>
</p>
```
:::

## Get started

If you haven't already, install the tidyverse (do this in the console, not in your script):

```{r}
#| eval: false
install.packages("tidyverse")
```

Make sure you have the tidyverse loaded at the top of your script:

```{r}
#| eval: false
library(tidyverse)
```

## Read a CSV file

1. Download the file below and save it in the `data/` folder of your project.
2. Use `read_csv()` to read the file into R and save it in a variable called `trees`.
3. Explore the data:
    - Use `summary(trees)` — how many tree species are there? What is the tallest tree?
    - Use `view(trees)` to look at the full table
    - Use `$` to access the `height_m` column and calculate the mean height

```{r}
#| echo: false
downloadthis::download_file(
  path = here::here("data/tree_growth.csv"),
  output_name = "tree_growth",
  button_label = "Download tree_growth.csv",
  button_type = "default",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

## Read an Excel file

1. Download the Excel file below and save it in the `data/` folder of your project.
2. Load the `readxl` package with `library(readxl)`.
3. Use `read_excel()` to read the file into R and save it in a variable.
4. Explore the data with `summary()`.

```{r}
#| echo: false
downloadthis::download_file(
  path = here::here("data/bird_observations.xlsx"),
  output_name = "bird_observations",
  button_label = "Download bird_observations.xlsx",
  button_type = "default",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

## Write data to a file

Take the `trees` tibble you just read in and write it to a new file:

1. Use `write_csv()` to save it as `trees_copy.csv` in your `data/` folder.
2. Check in the Files pane that the file was created.

## Challenge: a slightly messy file

1. Download the file below and save it in the `data/` folder.
2. Try reading it with `read_csv()`. Something is wrong — can you figure out what?
    - Hint: open the file in a text editor or in RStudio (File → Open File) to see its structure.
3. Use the appropriate argument of `read_csv()` to fix the problem. Check `?read_csv` for an argument that lets you skip lines at the top of a file.

```{r}
#| echo: false
downloadthis::download_file(
  path = here::here("data/water_quality.csv"),
  output_name = "water_quality",
  button_label = "Download water_quality.csv",
  button_type = "default",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

## Optional tasks (if you finish early)

You can do these in any order — or skip them and just take a break.

### Challenge: an even messier file

Download the file below and try to read it into R. This file has multiple problems — you'll need more than one argument to fix them.

:::{.callout-tip collapse="true"}

## Hint (unfold to see)

The file has metadata lines on top **and** uses a different delimiter (not a comma).

:::

```{r}
#| echo: false
downloadthis::download_file(
  path = here::here("data/soil_nutrients_messy.csv"),
  output_name = "soil_nutrients_messy",
  button_label = "Download soil_nutrients_messy.csv",
  button_type = "default",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

### Clean messy column headers

After reading the messy soil data, try using `janitor::clean_names()` on it. What does it do to the column names?

You may need to install the janitor package first: `install.packages("janitor")`.

### Read your own data

If you have your own research data, try reading it into R:

1. Copy a data file into the `data/` folder of your project
2. Use the appropriate `read_*()` function to read it in
3. Did it work? If not, check `?read_csv` or `?read_excel` for arguments that might help
