## Get started

We will again use the `penguins` data included in R.

As always, before you get started with the exercise, make sure that you load the `tidyverse`

```{r echo=TRUE, eval=FALSE}
library(tidyverse)
```

## Compare the flipper length of penguins using statistical tests

Question: Does the mean flipper length differ between the 3 penguin species?

Do separate comparisons for

- Gentoo vs. Adelie
- Gentoo vs. Chinstrap
- Adelie vs. Chinstrap

Before you start, create a subset of each species as a vector.

For this, you can use the `$` operator:

```{r eval=FALSE, echo = TRUE}
adelie <- filter(penguins, species == "Adelie")$flipper_len
chinstrap <- filter(penguins, species == "Chinstrap")$flipper_len
gentoo <- filter(penguins, species == "Gentoo")$flipper_len
```

Follow the decision tree for statistical tests:

```{r echo = FALSE}
knitr::include_graphics(here::here(
  "sessions/slides/img/day3/tests_overview.png"
))
```

After you have the test results, visualize the differences in flipper length between species.

You can e.g. choose one of the following plots:

- Boxplot with notches
- Mean and standard error of the mean (as mean with errorbar, as barplot, ...)

Add a `geom_signif()` layer to the plot you just created to indicate your test results

### For the fast ones

- Use a different visualization method than before to show the differences in flipper length between species
- Change the appearance of the plot (colors, themes, ...) from the task