---
title: "Task: Statistical tests"
number-sections: false
---

[← Back to session page](../10_tests.qmd)

:::{.callout-note}

## Find the solution here after the session ends.

```{=html}
<script>const course_day = 3;</script>
<script src="../../timer_solution.js"></script>
<p class="r_class_solution" style="display:none">
  <a href="../solutions/tests-solution.html" target="_blank">Click here for the solution.</a>
</p>
```
:::

## Get started

We will again use the `penguins` data included in R.

As always, before you get started with the exercise, make sure that you load the `tidyverse`

```{r}
#| eval: false
library(tidyverse)
```

## Compare the flipper length of penguins using statistical tests

Question: Does the mean flipper length differ between the 3 penguin species?

Do separate comparisons for

- Gentoo vs. Adelie
- Gentoo vs. Chinstrap
- Adelie vs. Chinstrap

Before you start, create a subset of each species as a vector.

For this, you can use the `$` operator:

```{r}
#| eval: false
adelie <- filter(penguins, species == "Adelie")$flipper_len
chinstrap <- filter(penguins, species == "Chinstrap")$flipper_len
gentoo <- filter(penguins, species == "Gentoo")$flipper_len
```

Follow the decision tree for statistical tests:

```{r}
#| echo: false
knitr::include_graphics(here::here(
  "sessions/slides/img/day3/tests_overview.png"
))
```

After you have the test results, visualize the differences in flipper length between species.

You can e.g. choose one of the following plots:

- Boxplot with notches
- Mean and standard error of the mean (as mean with errorbar, as barplot, ...)

Add a `geom_signif()` layer to the plot you just created to indicate your test results

## For the fast ones

You can do these in any order, or skip them and just take a break.

- Use a different visualization method than before to show the differences in flipper length between species
- Change the appearance of the plot (colors, themes, ...) from the task