---
title: "Data visualization with ggplot2"
subtitle: "Day 2 - Introduction to Data Analysis with R"
author: "Selina Baldauf"
institute: "Freie Universität Berlin - Theoretical Ecology"
date: today
date-format: long
format: 
  revealjs:
    footer: "Selina Baldauf // Data visualization with ggplot2"
    highlight-style: breeze
    mainfont: Cabinet Grotesk
    slide-number: true
    show-slide-number: all
    incremental: true
    self-contained: true
    code-line-numbers: false
    auto-stretch: false
    scrollable: false
    theme: slides.scss
    fig-align: center
    fig-width: 6.3
    fig-height: 6.3
execute: 
  echo: true
  warning: false
  message: false
  cache: true
knitr: 
  opts_chunk: 
    collapse: true
    comment: "#>"
from: markdown+emoji
filters:
  - roughnotation
  - flourish
---


```{r setup, include=FALSE}
library(tidyverse)
library(gapminder)
theme_set(theme_grey(base_size = 16))
```

# A ggplot showcase{.inverse}

> Example plots you can create with ggplot

## A ggplot showcase

:::{.r-stack}

![Visualization by [Jake Kaupp](https://twitter.com/jakekaupp), code available on [Github](https://github.com/jkaupp/tidytuesdays/tree/master/2021/week27)](https://raw.githubusercontent.com/jkaupp/tidytuesdays/master/2021/week27/tw27_plot.png)

:::

## A ggplot showcase

![Visualizations produced by the [BBC News data team](https://github.com/bbc/bbplot)](https://raw.githubusercontent.com/bbc/bbplot/master/chart_examples/bbplot_example_plots.png)

## A ggplot showcase

![Visualization by [Cédric Scherer](https://twitter.com/CedScherer), code available on [Github](https://github.com/z3tt/TidyTuesday/blob/master/R/2020_31_PalmerPenguins.Rmd)](https://raw.githubusercontent.com/Z3tt/TidyTuesday/master/plots/2020_31/2020_31_PalmerPenguins.png){width=35%}

## A ggplot showcase

![Overview of different plot possibilites created by Cédric Scherer. Check out [his tutorial](https://cedricscherer.netlify.app/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/#toc) to learn how to make all these plots with ggplot](img/day2/ggplot_showcase_cedric.png){width=60%}

## The ggplot2 package

- **Consistent** grammar/structure
- **Flexible** structure allows you to produce any type of plots
- Highly **customizable appearance** (themes)
- Many **extension packages** that provide additional plot types, themes, colors, animation, ...
  - See [here](https://exts.ggplot2.tidyverse.org/gallery/) for a list of ggplot extensions
- Active community that provides help and inspiration
- Perfect package for **exploratory data analysis** and **beautiful plots**

:::{.fragment}

Part of the tidyverse

```{r}
library(tidyverse)
# Or load it individually:
# library(ggplot2)
```

:::

## The data

The `gapminder` dataset contains data on life expectancy, GDP per capita, and population for 142 countries from 1952 to 2007.

```{r}
# install.packages("gapminder")
library(gapminder)
gapminder
```

. . .

How is life expectancy related to GDP per capita?

# Building a ggplot {.inverse}

> Step by step

## Step 1: `ggplot(data)`

The `ggplot()` function initializes a ggplot object. Every ggplot needs this function.

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(gapminder)
```

- Empty plot because we haven't told ggplot what to put on the axes yet

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(gapminder)
```

:::

:::

## Step 2: add `aes(x, y)`

The **aes**thetics define how data variables are mapped to plot properties.

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp
  )
)
```

- Axes are set up, but no data points yet
- Scales adapt automatically to the range of the data

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp
  )
)
```

:::

:::


## Step 3: add a `geom`

`geoms` define how data points are represented. There are many different geoms to chose from

![from [ggplot cheatsheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/data-visualization-2.1.pdf)](img/day2/geom_ggplot.png){width=60%}

## Step 3: add a `geom`

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp
  )
) +
  geom_point()
```

- New layers are added with `+`
- `data` and `aes` defined in `ggplot` call are inherited to all plot layers

:::{.fragment}
Does this depend on the continent?
:::

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp
  )
) +
  geom_point()
```

:::

:::

## Add color: `aes(color = ...)`

Map a variable to the **color** aesthetic to distinguish groups:

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|   - target: "color = continent"
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point()
```

- `color = continent` maps the continent to point color
- A legend is added automatically

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point()
```

:::

:::


## Other aesthetics: `size`

Besides `color`, you can also map variables to `size`:

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|  - target: "size = pop"
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent,
    size = pop
  )
) +
  geom_point()
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent,
    size = pop
  )
) +
  geom_point()
```

:::

:::

## Other aesthetics: `shape`

Besides `color`, you can also map variables to `shape`:

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|  - target: "shape = continent"
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent,
    shape = continent
  )
) +
  geom_point()
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent,
    shape = continent
  )
) +
  geom_point()
```

:::

:::

## Changing the scales of the aesthetics

The scales onto which the aesthetic elements are mapped can be changed.

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point()
```

- GDP is condensed on the left (few large values)
- A log scale would spread the data out

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point()
```

:::

:::

## `scale_x_log10`

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|  - target: "scale_x_log10()"
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  scale_x_log10()
```

- Scales can be changed for all elements of `aes`:

:::{.fragment}

<code><b>scale\_[aes-name]{.highlight-ylw}\_[scale-type]{.highlight-pink}</b></code>

Here we scale the <b>[x]{.highlight-ylw}</b> aesthetic to <b>[log10]{.highlight-pink}</b>.

:::

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  scale_x_log10()
```

:::

:::


## `geom_smooth`

Add a smoothing line that helps see patterns in the data

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|    - target: "geom_smooth(method = 'lm')"
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10()
```

- With `method = "lm"`, a linear regression line is added
- All geoms done separately for each continent because color is defined globally

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_log10()
```

:::

:::

## `geom_boxplot`

Compare a numeric variable across groups:

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp
  )
) +
  geom_boxplot()
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp
  )
) +
  geom_boxplot()
```

:::

:::

## `geom_boxplot`

Compare a numeric variable across groups:

:::{.columns}

:::{.column width="50%"}

```{r}
#| label: anatomy-of-boxplot
#| echo: false
set.seed(123)
data <- data.frame(
  value = c(rnorm(100, mean = 10, sd = 2)),
  group = "Example"
)
stats <- summary(data$value)
iqr <- IQR(data$value)
outlier_upper <- stats[5] + 1.5 * iqr
outlier_lower <- stats[2] - 1.5 * iqr
boxplot_anatomy <- ggplot(data, aes(x = group, y = value)) +
  geom_boxplot(width = 0.3, outlier.size = 3, size = 1) +
  annotate(
    "text",
    x = 1.4,
    y = stats[6],
    label = "Maximum (excluding outliers)",
    hjust = 0
  ) +
  annotate(
    "text",
    x = 1.4,
    y = stats[5],
    label = "75th percentile (Q3)",
    hjust = 0
  ) +
  annotate("text", x = 1.4, y = stats[3], label = "Median", hjust = 0) +
  annotate(
    "text",
    x = 1.4,
    y = stats[2],
    label = "25th percentile (Q1)",
    hjust = 0
  ) +
  annotate("text", x = 1.4, y = stats[1], label = "Outlier", hjust = 0) +
  annotate(
    "segment",
    x = 1.3,
    xend = 1.2,
    y = c(stats[1], stats[2], stats[3], stats[5], stats[6]),
    yend = c(stats[1], stats[2], stats[3], stats[5], stats[6])
  ) +
  annotate(
    "text",
    x = 0.5,
    y = mean(c(stats[2], stats[5])),
    label = "IQR\n(Inter Quartile\nRange)",
    hjust = 0.5
  ) +
  theme_void() +
  theme(
    axis.text.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.background = element_rect(fill = "transparent", color = NA),
    plot.background = element_rect(fill = "transparent", color = NA)
  ) +
  coord_cartesian(xlim = c(0.7, 2.5))
boxplot_anatomy
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp
  )
) +
  geom_boxplot()
```

:::

:::

## `geom_boxplot` with color

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|  - target: "color = continent"
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp,
    color = continent
  )
) +
  geom_boxplot()
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp,
    color = continent
  )
) +
  geom_boxplot()
```

:::

:::

## `geom_boxplot` with fill

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|     - target: "fill = continent"
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp,
    fill = continent
  )
) +
  geom_boxplot()
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp,
    fill = continent
  )
) +
  geom_boxplot()
```

:::

:::

## `geom_tile`

You can create a simple heatmap with `geom_tile`

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = year,
    y = continent,
    fill = lifeExp
  )
) +
  geom_tile()
```

- Here we would have to choose a different color scheme to see differences

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = year,
    y = continent,
    fill = lifeExp
  )
) +
  geom_tile()
```

:::

:::


## `geom_histogram`

See the distribution of a single numeric variable:

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(x = lifeExp)
) +
  geom_histogram()
```

- ggplot counts the observations for you (y-axis)

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(x = lifeExp)
) +
  geom_histogram()
```

:::

:::

## `geom_histogram` with groups

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = lifeExp,
    fill = continent
  )
) +
  geom_histogram(
    position = "identity",
    alpha = 0.5
  )
```

- By default histograms are stacked
- Use `position = "identity"` to overlap them
- `alpha` makes sure you see overlapping areas
- Here it's a bit too crowded to see differences between continents

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = lifeExp,
    fill = continent
  )
) +
  geom_histogram(
    position = "identity",
    alpha = 0.5
  )
```

:::

:::


## Small multiples with `facet_wrap`

Split your plots along one variable with `facet_wrap`

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|   - target: "facet_wrap(vars(continent))"
ggplot(
  gapminder,
  aes(
    x = lifeExp,
    fill = continent
  )
) +
  geom_histogram() +
  facet_wrap(vars(continent))
```

- Especially useful when color/shape gets crowded

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = lifeExp,
    fill = continent
  )
) +
  geom_histogram() +
  facet_wrap(vars(continent))
```

:::

:::


## Small multiples with `facet_grid`

Split your plots along two variables with `facet_grid`

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  scale_x_log10() +
  facet_grid(
    rows = vars(continent),
    cols = vars(year)
  )
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  scale_x_log10() +
  facet_grid(
    rows = vars(continent),
    cols = vars(year)
  ) +
  theme_gray(base_size = 14)
```

:::

:::

## So many other geoms

Check out:

:::{.nonincremental}

- [ggplot2 cheatsheet](https://ggplot2.tidyverse.org/#cheatsheet)
- [ggplot2 cookbook](https://r-graphics.org/)
- [ggplot2 tutorial](https://www.cedricscherer.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/) 

:::


## Summary: The ggplot skeleton

Every ggplot has the same basic structure:

```{r}
#| eval: false
ggplot(
  data,
  aes(x = ..., y = ..., color = ...)
) +
  geom_xxx()
```

:::{.nonincremental}

- `ggplot()`: initialize the plot with data and aesthetic mappings
- `aes()`: map variables to visual properties (x-axis, y-axis, color, shape, size, ...)
- `geom_xxx()`: define how data points are drawn
- Layers are connected with `+`

:::


# Now you {.inverse}

[Task 1 (45 min)]{.highlight-blue}<br>

[Exploratory data analysis with the penguin data set]{.big-text}

**Find the task description [here](https://selinazitrone.github.io/intro-r-data-analysis/sessions/tasks/ggplot-task-1.html)**

# Beautifying plots {.inverse}

> From default to publication-ready

## Mapping vs. setting aesthetics

. . .

:::{.columns}

:::{.column width="50%"}

Inside `aes()`: **map** a variable to color

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point()
```

```{r}
#| echo: false
#| fig-width: 3.5
#| fig-height: 3.3
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point(size = 1) +
  scale_x_log10() +
  theme_gray(base_size = 10)
```

:::

:::{.column width="50%"}

:::{.fragment}

Inside `geom`: **set** color of all points

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp
  )
) +
  geom_point(
    color = "steelblue"
  )
```

```{r}
#| echo: false
#| fig-width: 3.5
#| fig-height: 3.3
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp
  )
) +
  geom_point(
    color = "steelblue",
    size = 1
  ) +
  scale_x_log10() +
  theme_gray(base_size = 10)
```

:::

:::

:::


## Mapping vs. setting aesthetics

You can change point shape, size, and transparency in the same way:

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp
  )
) +
  geom_point(
    size = 3, # size
    shape = 17, # shape
    color = "steelblue", # color
    alpha = 0.5 # transparency (0-1)
  ) +
  scale_x_log10()
```

:::{.fragment}

```{r}
#| echo: false
#| fig-width: 6
#| fig-height: 2.1
shapes <- data.frame(
  shape = c(0:19, 22, 21, 24, 23, 20),
  x = 0:24 %/% 5,
  y = -(0:24 %% 5)
)
ggplot(shapes, aes(x, y)) +
  geom_point(aes(shape = shape), size = 5, fill = "red") +
  geom_text(aes(label = shape), hjust = 0, nudge_x = 0.15) +
  scale_shape_identity() +
  expand_limits(x = 2) +
  theme_void()
```

[Shape](http://www.cookbook-r.com/Graphs/Shapes_and_line_types/) and [color](https://derekogle.com/NCGraphing/resources/colors) codes

:::

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp
  )
) +
  geom_point(
    size = 3,
    shape = 17,
    color = "steelblue",
    alpha = 0.5
  ) +
  scale_x_log10()
```

:::

:::


## Mapping vs. setting aesthetics

Mapping and setting aesthetics can also be combined:

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
#| flourish:
#|  - target: "color = continent"
#|  - target: "size = 2"
#|  - target: "alpha = 0.6"
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point(
    size = 2,
    alpha = 0.6
  ) +
  scale_x_log10()
```

:::

:::{.column width="50%"}

```{r}
#| echo: false
ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point(
    size = 2,
    alpha = 0.6
  ) +
  scale_x_log10()
```

:::

:::


## Starting point

Let's take this plot and make it look better step by step:

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot <- ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  scale_x_log10()
my_plot
```

- Saving the plot in a variable `my_plot` so we can build on it
- Other plot layers can still be added to `my_plot`

:::

:::{.column width="50%"}

```{r}
#| echo: false
my_plot <- ggplot(
  gapminder,
  aes(
    x = gdpPercap,
    y = lifeExp,
    color = continent
  )
) +
  geom_point() +
  scale_x_log10()
my_plot
```

:::

:::

## `scale_color_manual()`: choose your own colors

Change the colors of the color aesthetic:

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot <- my_plot +
  scale_color_manual(
    values = c(
      "darkorange",
      "steelblue",
      "forestgreen",
      "orchid",
      "grey70"
    )
  )
my_plot
```

- One color per group: number of colors must match the number of groups
- Colors can be [names](https://derekogle.com/NCGraphing/resources/colors) or hex codes like `"#FF6B35"`

:::

:::{.column width="50%"}

```{r}
#| echo: false
my_plot <- my_plot +
  scale_color_manual(
    values = c(
      "darkorange",
      "steelblue",
      "forestgreen",
      "orchid",
      "grey50"
    )
  )
my_plot
```

:::

:::


## `scale_color_viridis_d`

Change the colors of the color aesthetic:

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot +
  scale_color_viridis_d(
    option = "magma"
  )
```

- The viridis color palette is designed for viewers with common forms of color blindness
- `viridis_d` for discrete colors, `viridis_c` for continuous colors
- Different options of viridis color palettes:
`"magma"`, `"inferno"`, `"plasma"`, `"viridis"`, `"cividis"`

:::

:::{.column width="50%"}

```{r}
#| echo: false
my_plot +
  scale_color_viridis_d(option = "magma")
```

:::

:::


## `scale_color_brewer()`: use a color palette

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot +
  scale_color_brewer(
    palette = "Set2"
  )
```

- Pre-defined palettes, many are colorblind-friendly
- Other options: `"Dark2"`, `"Paired"`, `"Pastel1"`, ...
- Check `?scale_color_brewer` for all options

:::

:::{.column width="50%"}

```{r}
#| echo: false
my_plot +
  scale_color_brewer(
    palette = "Set2"
  )
```

:::

:::


## `scale_fill_*` vs. `scale_color_*`

:::{.columns}

:::{.column width="50%"}

```{r fig.height=4}
#| flourish:
#| - target: "color"
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp,
    color = continent
  )
) +
  geom_boxplot() +
  scale_color_brewer(
    palette = "Dark2"
  )
```

:::

:::{.column width="50%"}

```{r fig.height=4}
#| flourish:
#| - target: "fill"
ggplot(
  gapminder,
  aes(
    x = continent,
    y = lifeExp,
    fill = continent
  )
) +
  geom_boxplot() +
  scale_fill_brewer(
    palette = "Dark2"
  )
```

:::

:::


## `labs`: Change axis and legend titles and add plot title


```{r}
#| output-location: slide
#| fig-width: 8
#| fig-height: 6
my_plot <- my_plot +
  labs(
    x = "GDP per capita [US$]",
    y = "Life expectancy [years]",
    color = "Continent",
    title = "Wealth and life expectancy",
    subtitle = "Higher GDP per capita is associated with longer life expectancy",
    caption = "Data from the gapminder package"
  )
my_plot
```


## `theme_*`: change appearance

`ggplot2` offers many pre-defined themes that we can apply to change the appearance of a plot.

. . .

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot +
  theme_classic()
```

```{r echo=FALSE, fig.height=5}
my_plot +
  theme_classic(base_size = 14)
```

:::

:::{.column width="50%"}

```{r}
#| eval: false
my_plot +
  theme_bw()
```

```{r echo=FALSE, fig.height= 5}
my_plot +
  theme_bw(base_size = 14)
```

:::

:::

## `theme_*`: change appearance

`ggplot2` offers many pre-defined themes that we can apply to change the appearance of a plot.

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot +
  theme_minimal()
```

```{r echo=FALSE, fig.height= 5}
my_plot +
  theme_minimal(base_size = 14)
```

:::

:::{.column width="50%"}

```{r}
#| eval: false
my_plot +
  theme_linedraw()
```

```{r echo=FALSE, fig.height= 5}
my_plot +
  theme_linedraw(base_size = 14)
```

:::

:::

## `theme_*`: change appearance

Since `ggplot2` v. 4.0.0, you can change overall color choices of a pre-defined theme.

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot +
  theme_bw(
    ink = "#BBBBBB",
    paper = "#333333"
  )
```

- `paper`: affects background elements
- `ink`: affects foreground elements (text, lines, points, ...)
- `accent`: affects elements that are used to highlight information (like `geom_smooth()` lines)


:::

:::{.column width="50%"}

```{r echo=FALSE, fig.height= 5}
my_plot +
  theme_bw(
    base_size = 14,
    ink = "#BBBBBB",
    paper = "#333333"
  )
```

:::

:::

## `theme_*`: change appearance

Since `ggplot2` v. 4.0.0, you can change overall color choices of a pre-defined theme.

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot +
  theme_minimal(
    paper = "cornsilk",
    ink = "navy"
  )
```

:::{.nonincremental}
- `paper`: affects background elements
- `ink`: affects foreground elements (text, lines, points, ...)
- `accent`: affects elements that are used to highlight information (like `geom_smooth()` lines)

:::


:::

:::{.column width="50%"}

```{r echo=FALSE, fig.height= 5}
my_plot +
  theme_minimal(
    base_size = 14,
    paper = "cornsilk",
    ink = "navy"
  )
```

:::

:::

## `theme()`: fine-tune individual elements

:::{.columns}

:::{.column width="50%"}

```{r}
#| eval: false
my_plot <- my_plot +
  theme_classic() +
  theme(
    # Move legend inside the plot
    legend.position = "inside",
    # Define the position with x and y coordinates (0-1)
    legend.position.inside = c(0.90, 0.15),
    # Make the title bold
    plot.title = element_text(face = "bold"),
    # Make axis titles larger and bold
    axis.title = element_text(size = 14, face = "bold")
  )
my_plot
```

:::{.fragment}

The basic functioning of theme elements is:

```{r eval=FALSE}
theme(
  element_name = element_function()
)
```

:::

:::

:::{.column width="50%"}

```{r}
#| echo: false
my_plot <- my_plot +
  theme_classic(base_size = 14) +
  theme(
    # Move legend inside the plot
    legend.position = "inside",
    # Define the position with x and y coordinates (0-1)
    legend.position.inside = c(0.95, 0.05),
    # Make the title bold
    plot.title = element_text(face = "bold"),
    # Make axis titles larger and bold
    axis.title = element_text(size = 14, face = "bold")
  )
my_plot
final_plot <- my_plot
```

:::

:::

## `theme()`: fine-tune individual elements

- Check `?theme` for all theme elements and options
- Search: "ggplot theme ..." (e.my_plot. "ggplot theme remove legend")

. . .

![Theme reference sheet by [Isabella Benabaye](https://isabella-b.com/blog/ggplot2-theme-elements-reference/)](https://isabella-b.com/files/ggplot2-theme-elements-reference-v2.png){width=70%}


## `theme_set()`: set global theme

You can set a global theme that will be applied to all ggplot objects in the current R session.

```{r eval=FALSE}
# Globally set theme_minimal as the default theme
theme_set(theme_minimal())
```

Add this to the beginning of your script.

. . .

You can also specify some defaults, e.g. the text size:

```{r eval=FALSE}
theme_set(
  theme_minimal(
    base_size = 16,
    paper = "cornsilk",
    ink = "navy"
  )
)
```

This is very practical if you want to achieve a consistent look, e.g. for a scientific journal.

## `ggsave()`

A ggplot object can be saved on disk in different formats.

Without specifications:

```{r eval=FALSE}
# save plot my_plot in img as my_plot.pdf
ggsave(filename = "img/my_plot.pdf", plot = my_plot)
# save plot my_plot in img as my_plot.png
ggsave(filename = "img/my_plot.png", plot = my_plot)
```

. . .

Or with specifications:

```{r eval=FALSE}
# save a plot named my_plot in the img directory under the name my_plot.png
# with width 16 cm and height 9 cm
ggsave(
  filename = "img/my_plot.png",
  plot = my_plot,
  width = 16,
  heigth = 9,
  units = "cm"
)
```

Have a look at `?ggsave` to see all options.


# Now you {.inverse}

[Task 2 (30 min)]{.highlight-blue}<br>

[Make your penguin plots more beautiful]{.big-text}

**Find the task description [here](https://selinazitrone.github.io/intro-r-data-analysis/sessions/07_ggplot.html#beautify-the-plots)**
