---
title: "Introduction to version control with Git"
subtitle: "Day 2: Branching, Merging and collaboration workflows"
author: "Selina Baldauf"
institute: "Freie Universität Berlin - Theoretical Ecology"
date: today
date-format: long
format: 
  revealjs:
    footer: "Branching, Merging and Collaboration with Git"
    highlight-style: breeze
    mainfont: Cabinet Grotesk
    slide-number: true
    show-slide-number: all
    incremental: true
    self-contained: true
    code-line-numbers: false
    auto-stretch: false
    scrollable: true
    fig-align: center
    theme: slides.scss
execute: 
  echo: true
  warning: false
  message: false
  collapse: true
from: markdown+emoji
---

# {.background-no-title .inverse}

![Artwork by [Allison Horst](https://twitter.com/allison_horst)](img/AllisonHorst/ggplot2_exploratory.png){widht=70%}

## `ggplot(data)`

The `ggplot()` function initializes a ggplot object. Every ggplot needs this function.

. . .

```{r}
#| output-location: column-fragment
#| fig-width: 6.3
#| fig-height: 6.3
library(ggplot2) # or library(tidyverse)
library(lterdatasampler)

ggplot(data = and_vertebrates)
```

- Empty plot because we did not specify the mapping of data variables


# Examples in light mode

> And a block quote

## Heading 2

### Heading 3

Some text with [some yellow]{.highlight-ylw} and with [some green]{.highlight-grn} and [some pink]{.highlight-pink} and [some blue]{.highlight-blue}

[And a test of ]{.circle-sketch-highlight} a circle highlight.

#### Heading 4

Here we have [a link](google.com) and some **bold** and **italic**

```r
library(ggplot2)
ggplot(penugins, aes(1:10,1:0)) +
  geom_point("test")
```

```{r}
head(iris)
```

Here we have some `inline code`, or even real code `mean(1:10)`

# Examples in dark mode {.inverse}

> And a block quote

## Heading 2 {.inverse}

### Heading 3

Some text with [some yellow]{.highlight-ylw} and with [some green]{.highlight-grn} and [some pink]{.highlight-pink} and [some blue]{.highlight-blue}


[And a test of ]{.circle-sketch-highlight} a circle highlight.

#### Heading 4

Here we have [a link](google.com) and some **bold** and **italic**

```r
library(ggplot2)
ggplot(penugins, aes(1:10,1:0)) +
  geom_point()
```

Here we have some `inline code`, or even real code `mean(1:10)`

# Examples in blue mode {.inverse}

> And a block quote

## Heading 2 {.inverse}

### Heading 3

Some text with [some yellow]{.highlight-ylw} and with [some green]{.highlight-grn} and [some pink]{.highlight-pink} and [some blue]{.highlight-blue}


[And a test of ]{.circle-sketch-highlight} a circle highlight.

#### Heading 4

Here we have [a link](google.com) and some **bold** and **italic**

```r
library(ggplot2)
ggplot(penugins, aes(1:10,1:0)) +
  geom_point()
```

Here we have some `inline code`, or even real code `mean(1:10)`


