---
title: "Tidy data with tidyr"
subtitle: "Day 3 - Introduction to Data Analysis with R"
author: "Selina Baldauf"
institute: "Freie Universität Berlin - Theoretical Ecology"
date: today
date-format: long
format: 
  revealjs:
    footer: "Selina Baldauf // Tidy data with tidyr"
    highlight-style: breeze
    mainfont: Cabinet Grotesk
    slide-number: true
    show-slide-number: all
    incremental: true
    self-contained: true
    code-line-numbers: false
    auto-stretch: false
    scrollable: true
    theme: slides.scss
    fig-align: center
execute: 
  echo: true
  warning: false
  message: false
  cache: true
knitr: 
  opts_chunk: 
    collapse: true
    comment: "#>"
from: markdown+emoji
---

```{r setup, include=FALSE}
library(tibble)
library(tidyr)
```

# What is tidy data?{.inverse}

## What is tidy data?

![Illustration from the [Openscapes](https://www.openscapes.org/) blog [*Tidy Data for reproducibility, efficiency, and collaboration*](https://www.openscapes.org/blog/2020/10/12/tidy-data/) by Julia Lowndes and Allison Horst](img/AllisonHorst/tidydata_1.jpg){width=85%}

## What is tidy data?

Let's look at some examples

:::{.columns}

:::{.column width="40%"}

**Tidy**
```{r echo=FALSE}
data.frame(id = 1:6,
           name = c("floof", "max", "cat", "donut", "merlin", "panda"),
           color = c("gray", "black", "orange", "gray", "black", "calico")) |> 
  knitr::kable()
```

:::

:::{.column width="60%"}

:::{.fragment}

**Non-tidy**
```{r echo=FALSE}
data.frame(floof = "gray",
           max = "black",
           cat = "orange",
           donut = "gray",
           merlin = "black",
           panda = "calico") |> 
  knitr::kable()
```  

<br>
```{r echo=FALSE}
data.frame(gray = c("floof", "donut"),
           black = c("max", "merlin"),
           orange = c("cat", ""),
           calico = c("panda", "")) |> 
  knitr::kable()
```  

:::

:::

:::

. . .

<br>

Sometimes ***raw data*** is non-tidy because its structure is optimized for data entry or viewing rather than analysis.

## Why tidy data?

The main advantages of **tidy** data is that the `tidyverse` packages are built to work with it.


![Illustration from the [Openscapes](https://www.openscapes.org/) blog [*Tidy Data for reproducibility, efficiency, and collaboration*](https://www.openscapes.org/blog/2020/10/12/tidy-data/) by Julia Lowndes and Allison Horst](img/AllisonHorst/tidydata_3.jpg){width=70%}

## Example

Let's go back to the city data set from earlier:


```{r echo=FALSE}
cities_v <- c("Istanbul", "Moscow", "London", "Saint Petersburg", "Berlin", "Madrid", "Kyiv", "Rome", "Bucharest", "Paris")
population <- c(15.1e6, 12.5e6, 9e6, 5.4e6, 3.8e6, 3.2e6, 3e6, 2.8e6, 2.2e6, 2.1e6)
area_km2 <- c(2576, 2561, 1572, 1439,891,604, 839, 1285, 228, 105 )

country <- c("Turkey", "Russia", "UK", "Russia", "Germany", "Spain",
                        "Ukraine", "Italy", "Romania", "France")
# tidy
cities_tbl <- tibble(city = cities_v,
           population = population,
           area_km2 = area_km2,
           country = country)
```

```{r}
cities_tbl
```

. . .

This already looks pretty tidy.

## Same data different format

```{r eval=FALSE}
cities_untidy
```


```{r echo=FALSE}
cities_untidy <- unite(cities_tbl, col = "location", c("country", "city")) %>%
  pivot_longer(c("population", "area_km2"), names_to = "type") %>%
  pivot_wider(names_from = "location", values_from = "value")
print(cities_untidy, width = 90)
```

**What's not tidy here?**

. . .

- Each row has multiple observation
- At the same time, each observation is split across multiple rows
- Country and city variable are split into multiple columns
- Country and city variable values are united to one value

. . .

Let's tidy this data using functions from the `tidyr` package!

## `pivot_longer()`

**One variable split into multiple column**s can be solved with `pivot_longer`

```{r echo=FALSE}
print(cities_untidy, width = 90)
```

## `pivot_longer()`

**One variable split into multiple column**s can be solved with `pivot_longer`

```{r}
step1 <- pivot_longer(
    cities_untidy,                         # the tibble
    cols = Turkey_Istanbul:France_Paris,   # the columns to pivot from:to
    names_to = "location",                 # name of the new column
    values_to = "value")                   # name of the value column
```

. . .

```{r echo=FALSE}
print(step1, n=4)
```

## `pivot_longer()`

**One variable split into multiple column**s can be solved with `pivot_longer`

```{r}
#| eval: false
step1 <- pivot_longer(
    cities_untidy,                         # the tibble
    cols = Turkey_Istanbul:France_Paris,   # the columns to pivot from:to
    names_to = "location",                 # name of the new column
    values_to = "value")                   # name of the value column
```

Another way to select the columns to pivot:

```{r}
#| eval: false
#| code-line-numbers: "3"
step1 <- pivot_longer(
    cities_untidy,           # the tibble
    cols = !type,            # All columns except type#<<        
    names_to = "location",   # name of the new column
    values_to = "value")     # name of the value column
```

## `separate_wider_delim()`

**Multiple variable values that are united into one** can be separated using `separate_wider_delim`

```{r echo=FALSE}
print(step1,n=2)
```

. . .

```{r}
step2 <- separate_wider_delim(
  step1,                        # the tibble
  location,                     # the column to separate
  delim = "_",                  # the separator
  names = c("country", "city")) # names of new columns
```

. . .

```{r echo=FALSE}
print(step2, n=2)
```

. . .

The opposite function exists as well and is called `unite`. Check out `?unite` for details.

## `pivot_wider()`

**One observation split into multiple rows** can solved with `pivot_wider`

```{r echo=FALSE}
print(step2,n=2)
```

. . .

```{r}
step3 <- pivot_wider(
  step2,                      # the tibble
  names_from = type,          # the variables
  values_from = value)        # the values
```

. . .

```{r echo=FALSE}
print(step3, n=5)
```

## All steps in 1

We can also use a pipe to do all these steps in one:

```{r}
cities_tidy <- cities_untidy |>
  pivot_longer(Turkey_Istanbul:France_Paris,
    names_to = "location",
    values_to = "values"
  ) |>
  separate(location,
    sep = "_",
    into = c("country", "city")
  ) |>
  pivot_wider(
    names_from = type,
    values_from = values
  )
```

# Now you {.inverse}

[Task (30 min)]{.highlight-blue}<br>

[Tidy data with tidyr]{.big-text}

**Find the task description [here](https://selinazitrone.github.io/intro-r-data-analysis/sessions/09_tidyr.html)**
