Create a new script for the code of this task.
Use comments in your code to take notes.

### 2.1 Type conversion

Look at the vectors below. What do you think will happen? 

```{r echo=TRUE, eval=FALSE}
v1 <- c(1, "a", 2, 3)
v2 <- c(TRUE, TRUE, 1, FALSE, 0)
v3 <- c(0, "23", 5, 7)
v4 <- c(4L, 6L, 23.5345)
```

Copy the vectors to your script and check their data types using `typeof`.

Try to explicitly convert the vectors to integer and to character. What happens?

### 2.2 Working with vectors

You have the following vectors:

- `species`: name of species
- `bodywt_kg`: body weight of the species in kg
- `brainwt_g`: brain weight of the species in g

```{r echo=TRUE, eval=FALSE}
species <- c(
  "MountainBeaver", "Cow", "GreyWolf", "Goat",
  "GuineaPig", "Diplodocus", "AsianElephant", "Donkey",
  "Horse", "PotarMonkey", "Cat", "Giraffe",
  "Gorilla", "Human", "AfricanElephant", "Triceratops",
  "RhesusMonkey", "Kangaroo", "GoldenHamster", "Mouse",
  "Rabbit", "Sheep", "Jaguar", "Chimpanzee",
  "Rat", "Brachiosaurus", "Mole", "Pig"
)

bodywt_kg <- c(
  1.4, 465, 36.3, 27.7, 1., 11700, 2547, 187.1,
  521, 10, 3.3, 529, 207, 62, 6654, 9400,
  6.8, 35, 0.1, 0.02, 2.5, 55.5, 100, 52.2,
  0.3, 87000, 0.1, 192
)

brainwt_g <- c(
  8.1, 423, 119.5, 115, 5.5, 50, 4603, 419, 655, 115, 25.6,
  680, 406, 1320, 5712, 70, 179, 56, 1, 0.4, 12.1, 175,
  157, 440, 1.9, 154.5, 3, 180
)
```

#### Tasks

- Check if the following animals are contained in the `species` vector:

```{r echo=TRUE, eval=FALSE}
animals_to_check <- c("Snail", "Goat", "Chimpanzee", "Rat", "Dragon", "Eagle")
```

- Convert brain weight from g to kg and save result it in a new vector `brainwt_kg`
- Calculate mean and standard deviation of the brain weight
- Which species have brain weight that is larger than the mean brain weight?
- What is the ratio of brain weight to body weight in percent? Calculate it and save the result in a new variable
- Are there any animals with a larger brain to body ratio than humans? If yes, which ones?
  - Hint: find the ratio for humans and save it in a variable first


#### Extras

- Round the ratio vector to 4 decimal places with the `round` function
  - have a look at `?round` to check out how to use it
- Which animal has the smallest brain to body ratio?
  - Hint: have a look at the `min` function
- Add the following three animals to the data vectors

```{r echo=TRUE, eval=FALSE}
species_new <- c("Eagle", "Snail", "Lion")
brainwt_g_new <- c(40, NA, 500)
bodywt_kg_new <- c(18, 0.01, 550)
```

 Now calculate the mean brain weight again. Can you explain what happens? Can you fix it?
 
 - hint: have a look at `?mean`
