Create a new script for the code of this task.
Use comments in your code to take notes.

## Type conversion

Look at the vectors below. What do you think will happen? 

```{r}
v1 <- c(1, "a", 2, 3)
v2 <- c(TRUE, TRUE, 1, FALSE, 0)
v3 <- c(0, "23", 5, 7)
v4 <- c(4L, 6L, 23.5345)
```

Check out the data types of the vectors using `typeof`.

Try to explicitly convert all 4 vectors to integer. What happens?

## Working with vectors

You have the following vectors:

- `species`: name of species
- `bodywt_kg`: body weight of the species in kg
- `brainwt_g`: brain weight of the species in kg

```{r}
species <- c(
  "MountainBeaver", "Cow", "GreyWolf", "Goat",
  "GuineaPig", "Diplodocus", "AsianElephant", "Donkey",
  "Horse", "PotarMonkey", "Cat", "Giraffe",
  "Gorilla", "Human", "AfricanElephant", "Triceratops",
  "RhesusMonkey", "Kangaroo", "GoldenHamster", "Mouse",
  "Rabbit", "Sheep", "Jaguar", "Chimpanzee",
  "Rat", "Brachiosaurus", "Mole", "Pig"
)

bodywt_kg <- c(
  1.4, 465, 36.3, 27.7, 1., 11700, 2547, 187.1,
  521, 10, 3.3, 529, 207, 62, 6654, 9400,
  6.8, 35, 0.1, 0.02, 2.5, 55.5, 100, 52.2,
  0.3, 87000, 0.1, 192
)

brainwt_g <- c(
  8.1, 423, 119.5, 115, 5.5, 50, 4603, 419, 655, 115, 25.6,
  680, 406, 1320, 5712, 70, 179, 56, 1, 0.4, 12.1, 175,
  157, 440, 1.9, 154.5, 3, 180
)
```




```{r eval=FALSE}
as.character()
typeof()
is.character()
c()
:
seq()
rep()
length()
%in%
```

- type conversion
- vectorized calculations: relational, ....

# References

Wolfram Research, "Sample Data: Animal Weights" from the Wolfram Data Repository (2016) https://doi.org/10.24097/wolfram.43616.data
