---
title: Linear models with R
editor: 
  markdown: 
    wrap: 72
---

Packages needed for linear model analysis and examples:

```{r}
#| message: false
#| warning: false
# Example data
library(tidyverse) # Data manipulation
library(performance) # Model diagnostics
library(rstatix) # More friendly version of Tukey test
```

# Intro

When you fit a linear model (an Anova is a type of linear model), you
need to give the `lm` function a formula that describes the relationship
between the response variable and the predictor variables in your data.
This formula should reflect your hypothesis.

The formula is written as `y ~ x1 + x2 + ... + xn`, where `y` is the
response variable and `x1`, `x2`, ..., `xn` are the predictor variables.

Depending on the type of response variable, you can use `lm` to perform
different linear model types:

-   Regression model: `lm(y ~ x1)` with `x1` as a continuous variable
-   ANOVA: `lm(y ~ x1)` with `x1` as categorical variables
-   ANCOVA: `lm(y ~ x1 + x2)` with `x1` as a categorical variable and
    `x2` as a continuous variable

All these formulas can also contain interactions between the predictor
variables. Interactions between predictor variables are written as
`x1:x2`.

# General workflow

To do a linear model analysis in R, you can follow these steps (see
below for an example):

1.  Explore the data and your hypothesis using a quick visualization
2.  Fit the linear model according to your hypothesis using `lm`
3.  Get a summary of the model using `summary`: Check out model
    coefficients and effect sizes
4.  Check whether your explanatory variables are significant using
    `drop1`
5.  Check model assumptions using diagnostic plots and checking
    functions
6.  Take the plot from step 1, make it look nice and add the model fit.

# Example 1: ANOVA

This example uses the `penguins` dataset that is included in R

-   **Research question:** How does the penguin species and sex affect
    penguin body mass?
-   **Hyphothesis:** The body mass of penguins is higher for male
    penguins and differs between species.

## Explore the data

```{r}
ggplot(penguins, aes(x = species, y = body_mass, fill = sex)) +
  geom_boxplot() +
  labs(
    title = "Penguin body mass by species",
    x = "Species",
    y = "Body mass (g)"
  )
```

## Fit the linear model

I want to fit a linear model to test my hypothesis.

```{r}
# 2. Fit the linear model
penguin_lm <- lm(
  body_mass ~ species + sex, # Model formula to test your hypothesis
  data = penguins # Tibble containing the variables
)
```

If we also wanted to test the interaction between `species` and `sex`,
we have to add the interaction term to the model formula:
`body_mass ~ species + sex + species:sex`.

## Get a summary of the model

With the `summary` function, we can get model coefficients and effect
sizes:

```{r}
# 3. Get a summary of the model
summary(penguin_lm)
```

The model estimates can be found in the `Coefficients` table (column
`Estimat Std`). In case of an ANOVA, the `Estimate` column contains the
mean body mass for each species. It is a bit complex to disentangle, but
the values in the rows mean:

-   `(Intercept)`: The mean body mass of the female penguins of the
    first species (Adelie penguin)
-   `speciesChinstrap`: The difference in body mass between female
    Adelie and Chinstrap penguins (Chinstrap = 3372 g + 27 g)
-   `speciesGentoo`: The difference in body mass between female Adelie
    and Gentoo penguins (Gentoo = 3372 g + 1377 g)
-   `sexmale`: The difference in body mass between female and male
    penguins

## Check whether the explanatory variable is significant

To check whether both `species` and sex have a significant effect on
body mass, we can use the `drop1` function with an F-test. With this
test, we check whether the model would become significantly worse if we
remove the predictors.

```{r}
# 4. Check whether the explanatory variable is significant
drop1(penguin_lm, test = "F")
```

The p-value in both the species and the sex column of the result is \<
0.05, which means that both species and sex are significant predictors
of body mass. If we would remove these terms from the model, then the
model would become significantly worse.

## Check model assumptions

Linear models have 3 main assumptions that need to be fulfilled:

-   normality of model residuals
-   constant variance of model residuals
-   no strong outliers

We need to check them for every model. If they are violated, then we
have to use other types of models

### Visual checks

We can the model assumptions using diagnostic plots:

```{r}
#| fig-width: 10
#| fig-height: 10
performance::check_model(penguin_lm)
```

On top of each of the plots is written what you should look for. The
data are never perfect, so small deviations are totally ok.

### Other tests

You can also perform the most important checks manually and the results
are printed in the console for you. They are also provided by the
\`performance\` package:

```{r}
# Check for constant variance
performance::check_heteroscedasticity(penguin_lm)
# Check for influential observations (outliers)
performance::check_outliers(penguin_lm)
# check normality of residuals
performance::check_normality(penguin_lm)
```

If all these tests are ok, you can be confident that your model is
valid.

## Posthoc test to compare groups

If the ANOVA is significant, we can perform a posthoc test to compare
the groups. For this, we can use the `tukey_hsd` function from the
`rstatix` package. It is a wrapper around the `TukeyHSD` function from
base R that makes the output more friendly. This function performs
pairwise comparisons between the groups and corrects for multiple
testing.

```{r}
# Posthoc test to compare groups
tukey_hsd(penguin_lm)
```

# Further reading

-   [Book chapter on linear
    modelling](https://intro2r.com/simple_lm.html)
