---
editor_options:
  chunk_output_type: console
output: pdf_document
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(EcoData)
```

```{r packages}
library(tidyverse)
```

```{r get_data}
birds <- as_tibble(birdabundance)
#write_csv(birds, file = "./data/birdabundance.csv")
```

# Look at the structure of the data

What are the variables? What are the values?

First look at the summary and structure of the data

- only numeric columns
- response column is ABUND
- Summary shows not missing values

```{r}
str(birds)
summary(birds)
```

# Step 0: organize data

First, I want to rename some of the columns to make the data set easier to work with:

- Change all column headers to lower case
- rename column `yr.isol` to `isol_since`
- add a new column with years since isolation
- change `graze` variable to a factor

```{r}
birds <- birds %>% 
  rename_with(tolower, everything()) %>% 
  rename(isol_since = yr.isol) %>% 
  mutate(isol_years = 2021 - isol_since) %>% 
  mutate(graze = factor(graze))
```

# Step 1: Exploratory plotting

I will start with some exploratory data analysis using `ggplot2`.

#### Question: Which factors influence bird abundance most?

#### Isolation time and grazing

Looks like:
-  the longer a site is isolated the higher the grazing intensity is
-  the longer a site is isolated the lower the abundance
-  grazing intensity does not affect abundance but years of isolation do


```{r}
ggplot(birds, aes(x = isol_years, y = abund, color = graze)) +
  geom_point()
```

I there an interaction between grazing intensity and time since isolation?

Does not look like it. But the sites with a very high grazing intensity seem to be isolated since a very long time.

```{r}
ggplot(birds, aes(x = graze, y = isol_years)) +
  geom_boxplot() +
  geom_point(position = position_jitter(seed = 123, width = 0.2), alpha = 0.4, color = "cyan4")
```


#### Distance to forest

Does the bird abundance depend on the distance to the nearest forest patch?

Does not seem like it. There is no clear pattern in showing that the distance to the nearest forest 
affects bird abundance. However, the distances are all quite small considering the radius of bird movements

```{r}
ggplot(birds, aes(x = dist, y = abund)) +
  geom_point()+
  geom_smooth(method="lm")
```

How about the distance to the nearest large forest patch?

Also here, there does not seem to be a clear pattern

```{r}
ggplot(birds, aes(x=ldist, y=abund))+
  geom_point()+
  geom_smooth(method = "lm")
```

#### Altitude

Is there an effect of altitude?

It seems like the higher the altitude, the higher the bird abundance.

```{r}
ggplot(birds, aes(x = alt, y = abund)) +
  geom_point(aes(color = graze)) +
  geom_smooth(method = "lm")
```

#### Area

Is there an effect of the area of the nearest fragment?

```{r}
ggplot(birds, aes(x=area, y = abund))+
  geom_point()
```

Looks like a log transformation of area could help

```{r}
ggplot(birds, aes(x = area, y = abund, color = graze)) +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method = "lm", se = FALSE)
```

Looks like there is a clear relationship here.

# Step 2: Some statistical tests and models

First test a model with interaction between grazing intensity and area of the forest fragement.

```{r}
lm1 <- lm(abund ~ graze * area, data = birds)
drop1(lm1, test = "F")
```

There is a significant interaction between the grazing intensity and the area of the remaining forest fragment.

But are the assumptions of a linear model fulfilled?

```{r}
performance::check_model(lm1)
```

It looks like the normality of residuals is not fulfilled. 
Looking back at the plot from before, it might make sense to test a model with log-transformed area and a sqrt transfomed abundance.

(Sometimes the square root transformation can help with count data).

```{r}
lm1b <- lm(abund~graze*log(area), data = birds)
drop1(lm1b, test = "F")
# update model without interaction
lm1c <- lm(sqrt(abund)~graze + log(area), data = birds)
drop1(lm1c, test = "F")
```

```{r}
performance::check_model(lm1c)
```

This looks much better. However, it might also be a good idea to use a poisson glm in this case

```{r}
glm1 <- glm(abund ~ graze + log(area), data  =birds, family = "poisson")
drop1(glm1, test = "Chisq")
```

```{r}
ggplot(birds, aes(x = area, y = abund, color = graze)) +
  geom_point() +
  scale_x_log10() +
  geom_smooth(method = "glm", se = FALSE, method.args = list(family = "poisson"))
```

Or using the predict function:

```{r}
pred_dat <- expand_grid(
  graze = factor(1:5),
  area = min(birds$area):max(birds$area)
)
pred_dat$abund <- predict(glm1, newdata = pred_dat, type = "response")

ggplot(birds, aes(x = area, y = abund, color = graze)) +
  geom_point() +
  scale_x_log10() +
  geom_line(data = pred_dat)
```

#### Correlation between variables

Are there correlated variables?

```{r}
birds %>% 
  select(area, dist, ldist, isol_years, alt) %>% 
  cor()
```

Use the `ggpairs` function from the `GGally` package to plot a matrix plot of all 
variables from the data and look at possible correlations.

```{r}
GGally::ggpairs(birds)
```

To get a better overview, I now select some independent variables that I am interested in:

```{r}
birds %>% 
  select(-isol_since, -site) %>% 
  GGally::ggpairs()
```


```{r}
ggplot(data = birds,
mapping = aes(
x = dist,
y = isol_years,
size = abund,
color = abund)) +
geom_point() +
scale_color_gradient(low="blue",high="red")+
scale_x_log10()
```


Including some categorical variables for exploratory plotting

```{r}
q_10 <- quantile(birds$area, 0.1)
q_25 <- quantile(birds$area, 0.25)
q_80 <- quantile(birds$area, 0.8)

birds %>% 
  mutate(
    area_class = case_when(
      area < q_10  ~ "tiny",
      between(area, q_10, q_25) ~ "small",
      between(area, q_25, q_80) ~ "medium",
      area > q_80 ~ "large"
    )
  )
```



