### 1.1 Getting started


```{r echo=TRUE, eval=TRUE, message=FALSE}
#install.packages("performance")

library(tidyverse)
library(performance)
```

### 1.2 Linear regression

It looks like the bill depth increases with increasing bill length. 

```{r, echo=TRUE}
g <- ggplot(penguins, aes(x = bill_len, y = bill_dep)) +
  geom_point()
g
```

- Fit a linear model to test this relationship
- Check for significant effects

The result of `drop1` shows that if we remove the bill length from the model, we significantly increase the RSS. The model with bill length is the better model and we can assume a significant relationship between bill lenght and bill depth.

```{r, echo=TRUE}
lm1 <- lm(bill_dep ~ bill_len, data = penguins)
drop1(lm1, test = "F")
```

- Check the model assumptions using diagnostic plots

```{r, echo=TRUE, fig.height=7}
performance::check_model(lm1)
```

The diagnostic plots do not look perfect. The reason for this is probably that we did not consider all variable that we should have in the model. Maybe by adding the species as a predictor, we can improve the diagnostic plots?

- Can you add the model result to the scatter plot?

```{r, echo=TRUE}
# Option 1: geom_smooth
g +
  geom_smooth(method = "lm", se = FALSE)
```

```{r, echo=TRUE}
# Option 2: extract coefficients
intercept <- lm1$coefficients[1]
slope <- lm1$coefficients[2]
g +
  geom_abline(slope = slope, intercept = intercept)
```

```{r, echo=TRUE}
# Option 3: using predict (this is a bit overkill for this simple example)
# Create data to predict from
pred_dat <- tibble(
  bill_len = seq(
    from = min(penguins$bill_len, na.rm = TRUE),
    to = max(penguins$bill_len, na.rm = TRUE)
  )
)
# Predict the response and add it to pred_data
pred_dat$bill_dep <- predict(lm1, newdata = pred_dat)
# Add a line with the new predicted data
g +
  geom_line(data = pred_dat, color = "cyan4")

```

### 1.3 Analysis of covariance

Does the relationship between bill length and bill depth differ between penguin species?

```{r, echo=TRUE}
g2 <- ggplot(penguins, aes(x = bill_len, 
                     y = bill_dep,
                     color = species)) +
  geom_point()
g2
```

Now fit the linear model again, this time including species as explanatory variable.

```{r, echo=TRUE}
# Without interaction
lm2a <- lm(bill_dep ~ bill_len + species, data = penguins)
# Without interaction
lm2b <- lm(bill_dep ~ bill_len * species, data = penguins)
```

- Test model assumptions for both models.

Both models don't violate the assumptions for linear models and the diagnostic plots look much better than the one for the model without species.

Note that the diagnostic plots show a high collinearity between explanatory variables. This makes sense because of course the interaction between species and bill length is highly correlated with both species and bill length. However, in the case of this interaction it is not a problem, so we can ignore this.

```{r,echo=TRUE, fig.height=7}
check_model(lm2a)
check_model(lm2b)
```

- Check for significant effects in both models

```{r, echo=TRUE}
drop1(lm2a, test = "F")
drop1(lm2b, test = "F")
```

I will choose to plot the simpler model without the interaction because the interaction between species and bill length is not significant. Generally, a model should be as simple as possible (of course without loosing explanatory power).

To plot the model with interaction I could conveniently use `geom_smooth`

```{r, echo=TRUE}
g2 +
  geom_smooth(method = "lm", se = FALSE)
```

To plot th model without interaction, you can either extract the model coefficients or use the predict function.

Extracting the model coefficients starts to become complicated, as we have 3 intercepts and 1 slope to extract:

```{r, echo=TRUE}
# Option 1: extract model coefficients
intercept_adelie <- lm2a$coefficients[1]
slope <- lm2a$coefficients[2]
intercept_chinstrap <- intercept_adelie + lm2a$coefficients[3]
intercept_gentoo <- intercept_adelie + lm2a$coefficients[4]

# lets add some custom colors
cols <- c("darkorange", "cyan4", "purple")

g2 +
  geom_abline(slope = slope, intercept = intercept_adelie, color = cols[1]) +
  geom_abline(slope = slope, intercept = intercept_chinstrap, color = cols[2]) +
  geom_abline(slope = slope, intercept = intercept_gentoo, color = cols[3]) +
  scale_color_manual(values = cols)
```
It is easier to use the `predict` function in this case:

```{r, echo=TRUE}
pred_dat <- expand_grid(
  bill_len = min(penguins$bill_len, na.rm = TRUE):max(penguins$bill_len, na.rm = TRUE),
  species = c("Adelie", "Chinstrap", "Gentoo")
)
pred_dat$bill_dep <- predict(lm2a, newdata = pred_dat)

g2 +
  geom_line(data = pred_dat) +
  ggsci::scale_color_futurama()
```

### 1.4 Anova

Do male and female penguins differ in weight and does this depend on the species?

First, we create a subset without penguins for which the sex is unknown

```{r, echo=TRUE}
penguins_sex <- filter(penguins, !is.na(sex))
```

The question can be represented for example with a boxplot

```{r, echo=TRUE}
ggplot(penguins_sex, aes(x = species, y = body_mass, fill = sex)) +
  geom_boxplot(notch = TRUE)
```

- Fit a linear model that represents this question

```{r, echo=TRUE}
# Without interaction
lm3a <- lm(body_mass ~ sex + species, data = penguins_sex)
# Without interaction
lm3b <- lm(body_mass ~ sex + species + sex:species, data = penguins_sex)
```

- Test for significant predictors

Both the single effect and the interaction is significant.

```{r, echo=TRUE}
drop1(lm3a, test = "F")
drop1(lm3b, test = "F")
```
Both look fine. Model can be used.

```{r, fig.height=7, echo=TRUE}
check_model(lm3a)
check_model(lm3b)
```

#### Extra

Some other plots with which we can represent the results of the Anova.

```{r, echo=TRUE}
penguins_sex %>%
  ggplot(aes(
    x = species,
    y = body_mass,
    color = sex
  )) +
  geom_boxplot() +
  geom_point(
    size = 2, alpha = 0.5,
    position = position_jitterdodge(
      seed = 123
    )
  ) +
  coord_flip() +
  ggsci::scale_color_d3() +
  labs(y = "Body mass [g]", x = "Species") +
  theme_bw() +
  theme(legend.position = c(0.85, 0.15))
```



```{r echo=TRUE}
ggplot(penguins_sex, aes(x = species, y = body_mass, color = sex)) +
  stat_summary(
    position = position_dodge(width = 0.5)
  ) +
  scale_color_manual(values = c("#00AFBB", "#E7B800")) +
  labs(y = "Body mass [g]")
```

In the next plot, you need to make sure that the `width` of the `position_dodge` is
the same for both the bar and the errorbar.

```{r echo=TRUE}
ggplot(penguins_sex, aes(x = species, y = body_mass, fill = sex)) +
  stat_summary(
    fun.y = mean,
    geom = "bar",
    width = 0.7,
    position = position_dodge(width = 0.8)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.8),
    width = 0.3
  ) +
  scale_fill_manual(values = c("#00AFBB", "#E7B800")) +
  labs(y = "Body mass [g]")
```

