---
title: "Solution to dplyr tasks"
code-line-numbers: false
---

## Get started

```{r echo=TRUE}
#| message: false
library(tidyverse)

```
```{r echo=FALSE}
theme_set(theme_grey())
```

## Data transformation with dplyr

**Find all penguins that ...**

1. ... have a bill length between 40 and 45 mm.

```{r echo=TRUE}
filter(penguins, between(bill_len, 40, 45))
# same as 
# filter(penguins, bill_len < 45 & bill_len > 40)
```

2. ... for which we know the sex.

```{r echo=TRUE}
filter(penguins, !is.na(sex))
```

3. ... which are of the species Adelie or Gentoo

```{r echo=TRUE}
filter(penguins, species %in% c("Adelie", "Gentoo")
# or
# filter(penguins, (species == "Adelie" | species == "Gentoo"))
```

4. ... lived on the island Dream in the year 2007. How many of them were from each of the 3 species?

```{r echo=TRUE}
filter(penguins, island == "Dream" & year == 2007) |> 
  count(species)
```

**Count ...**

1. ... the number of penguins on each island.

```{r echo=TRUE}
count(penguins, island)
```

2. ... the number of penguins of each species on each island.

```{r echo=TRUE}
count(penguins, island, species)
```

**Select ...**

1. ... only the variables species, sex and year

```{r echo=TRUE}
select(penguins, species, sex, year)
```

2. ... only columns that contain measurements in mm

```{r echo=TRUE}
select(penguins, ends_with("mm"))
# or
#select(penguins, contains("_mm"))
```

**Add a column ...**

1. ... with the ratio of bill length to bill depth

```{r echo=TRUE}
mutate(penguins,
       ratio = bill_len / bill_dep)
```

2. ... with abbreviations for the species (Adelie = A, Gentoo = G, Chinstrap = C).

```{r echo=TRUE}
mutate(penguins,
       species_short = case_when(
         species == "Adelie" ~ "A",
         species == "Gentoo" ~ "G",
         species == "Chinstrap" ~ "C"
       ))
```

**Calculate ...**

1. ... mean flipper length and body mass for the 3 species and male and female penguins separately

```{r echo=TRUE}
penguins |>
  summarize(
    mean_flipper = mean(flipper_len, na.rm = TRUE),
    mean_body = mean(body_mass, na.rm = TRUE),
    .by = c(species, sex)
  )  
```

2. Can you do the same but remove the penguins for which we don't know the sex first?

```{r echo=TRUE}
penguins |> 
  filter(!is.na(sex)) |> 
  summarize(
    mean_flipper = mean(flipper_len, na.rm = TRUE),
    mean_body = mean(body_mass, na.rm = TRUE),
    .by = c(species, sex)
  )
```

## Extras

1. Make a boxplot of penguin body mass with sex on the y-axis and facets for the different species. Can you remove the penguins with missing values for sex first?

```{r echo=TRUE}
penguins |>
  filter(!is.na(sex)) |>
  ggplot(aes(x = sex, y = body_mass)) +
  geom_boxplot() +
  facet_wrap(~species)
```

2. Make a scatterplot with the ratio of bill length to bill depth on the y axis and flipper length on the x axis? Can you distinguish the point between male and female penguins and remove penguins with unknown sex before making the plot?

```{r echo=TRUE}
penguins |>
  mutate(ratio = bill_len / bill_dep) |>
  filter(!is.na(sex)) |>
  ggplot(aes(x = flipper_len, y = ratio, color = sex)) +
  geom_point() +
  scale_color_manual(values = c("cyan4", "darkorange")) +
  labs(
    x = "Flipper lenght (mm)",
    y = "Ratio bill length / bill depth (-)"
  ) +
  theme_minimal()
```
