---
title: "Solution: Summarize and visualize"
code-line-numbers: false
---

## Get started

```{r echo=TRUE}
#| message: false
library(tidyverse)
```
```{r echo=FALSE}
theme_set(theme_grey())
penguins <- as_tibble(penguins)
```

## Count

1. Count the number of penguins on each island.

```{r echo=TRUE}
count(penguins, island)
```

2. Count the number of penguins of each species on each island.

```{r echo=TRUE}
count(penguins, island, species)
```

## Summarize

3. Calculate mean flipper length and body mass for the 3 species separately.

```{r echo=TRUE}
penguins |>
  summarize(
    mean_flipper = mean(flipper_len, na.rm = TRUE),
    mean_body = mean(body_mass, na.rm = TRUE),
    .by = species
  )
```

4. Calculate mean flipper length and body mass by species and sex. Remove penguins with unknown sex first.

```{r echo=TRUE}
penguins |>
  drop_na(sex) |>
  summarize(
    mean_flipper = mean(flipper_len, na.rm = TRUE),
    mean_body = mean(body_mass, na.rm = TRUE),
    .by = c(species, sex)
  )
```

## Combine dplyr and ggplot

5. Remove penguins with missing `sex`, then pipe into a boxplot of body mass with `sex` on the x-axis.

```{r echo=TRUE}
penguins |>
  drop_na(sex) |>
  ggplot(aes(x = sex, y = body_mass)) +
  geom_boxplot()
```

6. Remove penguins with missing `sex`, then make a scatterplot of bill length vs. bill depth, colored by species.

```{r echo=TRUE}
penguins |>
  drop_na(sex) |>
  ggplot(aes(x = bill_len, y = bill_dep, color = species)) +
  geom_point()
```

## For the fast ones

Summarize mean body mass by species, then pipe the result into a bar chart with `geom_col()`.

```{r echo=TRUE}
penguins |>
  summarize(
    mean_body = mean(body_mass, na.rm = TRUE),
    .by = species
  ) |>
  ggplot(aes(x = species, y = mean_body)) +
  geom_col()
```

Calculate the min, max, and mean flipper length per species.

```{r echo=TRUE}
penguins |>
  summarize(
    min_flipper = min(flipper_len, na.rm = TRUE),
    max_flipper = max(flipper_len, na.rm = TRUE),
    mean_flipper = mean(flipper_len, na.rm = TRUE),
    .by = species
  )
```

Sort the result by mean flipper length using `arrange()`.

```{r echo=TRUE}
penguins |>
  summarize(
    min_flipper = min(flipper_len, na.rm = TRUE),
    max_flipper = max(flipper_len, na.rm = TRUE),
    mean_flipper = mean(flipper_len, na.rm = TRUE),
    .by = species
  ) |>
  arrange(mean_flipper)
```
