---
title: "Solution: Filter, select, and mutate"
code-line-numbers: false
---

## Get started

```{r echo=TRUE}
#| message: false
library(tidyverse)
```
```{r echo=FALSE}
penguins <- as_tibble(penguins)
```

## Filter penguins

**Find all penguins that ...**

1. ... have a bill length between 40 and 45 mm.

```{r echo=TRUE}
filter(penguins, bill_len >= 40 & bill_len <= 45)
```

2. ... are of the species Adelie or Gentoo.

```{r echo=TRUE}
filter(penguins, species %in% c("Adelie", "Gentoo"))
# or
# filter(penguins, species == "Adelie" | species == "Gentoo")
```

3. ... lived on the island Dream in the year 2007.

```{r echo=TRUE}
filter(penguins, island == "Dream" & year == 2007)
```

## Remove missing values

4. Remove all penguins with missing values for `sex`.

```{r echo=TRUE}
drop_na(penguins, sex)
```

## Select columns

5. Select only the variables `species`, `sex`, and `year`.

```{r echo=TRUE}
select(penguins, species, sex, year)
```

6. Select only columns that start with `"bill"`.

```{r echo=TRUE}
select(penguins, starts_with("bill"))
```

## Add new columns

7. Add a column with the ratio of bill length to bill depth.

```{r echo=TRUE}
mutate(penguins, ratio = bill_len / bill_dep)
```

8. Add a column with abbreviations for the species (Adelie = A, Gentoo = G, Chinstrap = C).

```{r echo=TRUE}
mutate(
  penguins,
  species_short = case_when(
    species == "Adelie" ~ "A",
    species == "Gentoo" ~ "G",
    species == "Chinstrap" ~ "C",
    .default = NA
  )
)
```

## Combine with the pipe

9. Use the pipe to: remove rows with missing `sex`, keep only Adelie penguins, and select `species`, `sex`, and `body_mass`.

```{r echo=TRUE}
penguins |>
  drop_na(sex) |>
  filter(species == "Adelie") |>
  select(species, sex, body_mass)
```

## For the fast ones

Use `filter_out()` to exclude penguins from Torgersen island, then select only `species`, `island`, and `flipper_len`.

```{r echo=TRUE}
penguins |>
  filter_out(island == "Torgersen") |>
  select(species, island, flipper_len)
```

Create a `size_category` column with `case_when` based on body mass, in a pipe that also removes NAs and selects only `species`, `body_mass`, and `size_category`.

```{r echo=TRUE}
penguins |>
  drop_na(body_mass) |>
  mutate(
    size_category = case_when(
      body_mass < 3500 ~ "small",
      body_mass < 5000 ~ "medium",
      body_mass >= 5000 ~ "large"
    )
  ) |>
  select(species, body_mass, size_category)
```
