---
title: "Solution to ggplot Task 2: Beautify and save plots"
code-line-numbers: false
---

## Get started

First we need to load the `tidyverse` packages:

```{r eval=TRUE, echo=TRUE}
#| message: false
# install.packages("tidyverse")
library(tidyverse)
```

## Add labels

Here is an example of adding labels to the scatterplot from Task 1:

```{r}
#| message: false
#| warning: false
ggplot(penguins, aes(
  x = bill_len,
  y = bill_dep,
  color = species
)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(
    title = "Penguin bill dimensions",
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    color = "Species"
  )
```

## Change the colors

You can use `scale_color_manual()` to set custom colors:

```{r}
#| message: false
#| warning: false
ggplot(penguins, aes(
  x = bill_len,
  y = bill_dep,
  color = species
)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  labs(
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    color = "Species"
  )
```

## Apply a theme

Try different themes to change the overall look of the plot:

```{r}
#| message: false
#| warning: false
ggplot(penguins, aes(
  x = bill_len,
  y = bill_dep,
  color = species
)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  labs(
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    color = "Species"
  ) +
  theme_minimal()
```

## Reproduce the plot

The following code is adapted from the [palmerpenguins package website](https://allisonhorst.github.io/palmerpenguins/articles/examples.html).

```{r}
#| warning: false
#| message: false
ggplot(
  data = penguins,
  aes(
    x = bill_len,
    y = bill_dep,
    color = species,
    shape = species
  )
) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  labs(
    title = "Penguin bill dimensions",
    subtitle = "Bill length and depth for Adelie, Chinstrap and
    Gentoo Penguins at Palmer Station LTER",
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    color = "Penguin species",
    shape = "Penguin species"
  ) +
  theme_minimal() +
  theme(
    legend.position = c(0.85, 0.15),
    legend.background = element_rect(fill = "white", color = NA)
  )
```

What was changed compared to the basic plot?

- Make points larger and slightly transparent by setting `size` and `alpha` for the point layer
- Change to custom color scale
- Add title and subtitle with `labs`
- Change title of x-axis, y-axis and legend for color and shape aesthetic with `labs`
- Use `theme_minimal()` instead of default theme
- Change legend position to bottom right corner within the plot
  - Positions are relative to the bottom left corner of the plot
  - 0.85 (85% of plot width) to the right
  - 0.15 (15% of plot height) towards the top

## Save your plot

Example with one of the plots from above:

```{r echo = TRUE}
# First save the plot in a variable
flipper_box <- ggplot(penguins, aes(species, flipper_len, color = species)) +
  geom_boxplot(width = 0.3) +
  geom_jitter(alpha = 0.5, position = position_jitter(width = 0.2, seed = 123)) +
  ggsci::scale_color_uchicago() +
  labs(x = "Species", y = "Flipper length (mm)") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r echo= TRUE, eval=FALSE}
# save as png in /img directory of the project
ggsave(filename = "img/flipper_box.png", plot = flipper_box)
# save as pdf in /img directory of the project
ggsave(filename = "img/flipper_box.pdf", plot = flipper_box)
```

## For the fast ones

### Beautified histogram

```{r echo = TRUE}
ggplot(penguins, aes(x = flipper_len, fill = species)) +
  geom_histogram(alpha = 0.6) +
  ggsci::scale_fill_d3() +
  labs(
    y = "Frequency",
    x = "Flipper length [mm]",
    fill = "Penguin species"
  ) +
  theme_minimal()
```

### Beautified heatmap

```{r echo=TRUE}
ggplot(penguins, aes(
  x = species,
  y = sex,
  fill = flipper_len
)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_classic()
```

### The `patchwork` package

With the `patchwork` package, you can combine multiple ggplots into one plot. The package allows you to add annotations to the plot and to control the layout and appearance.

Below you find a simple example of two different penguin scatterplots. For more explanation and an overview of what is possible with the package, please have a look at the [package documentation](https://patchwork.data-imaginist.com/)

```{r echo=TRUE}
#| warning: false
library(patchwork)
plot_1 <- ggplot(penguins, aes(
  x = bill_len, y = bill_dep,
  color = species
)) +
  geom_point()

plot_2 <- ggplot(penguins, aes(
  x = bill_len, y = body_mass,
  color = species
)) +
  geom_point()

# Simple combination of 2 plots in patchwork
plot_1 + plot_2
```

And a more complex example where shared layers are defined for both plots:

```{r}
#| warning: false
# more complex combination with annotation and definition of shared layers
final_plot <- plot_1 + plot_2 +
  plot_layout(guides = "collect") +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") &
  theme_minimal() &
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) &
  labs(
    color = "Penguin species"
  ) &
  theme(
    plot.tag.position = c(0.1, 0.95),
    plot.tag = element_text(face = "bold")
  )

final_plot
```

## References

Check out the [package website](https://allisonhorst.github.io/palmerpenguins/articles/examples.html) of the `palmerpenguins` package. They have more nice examples of data visualizations that you can do with ggplot.

Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0. https://allisonhorst.github.io/palmerpenguins/. doi: 10.5281/zenodo.3960218.
