---
title: "Solution to working with vectors"
---

```{r}
#| include: false
#| eval: true
options(scipen = 999)
```

You have the following three vectors:

```{r echo=TRUE}
species <- c(
  "MountainBeaver", "Cow", "GreyWolf", "Goat",
  "GuineaPig", "Diplodocus", "AsianElephant", "Donkey",
  "Horse", "PotarMonkey", "Cat", "Giraffe",
  "Gorilla", "Human", "AfricanElephant", "Triceratops",
  "RhesusMonkey", "Kangaroo", "GoldenHamster", "Mouse",
  "Rabbit", "Sheep", "Jaguar", "Chimpanzee",
  "Rat", "Brachiosaurus", "Mole", "Pig"
)

bodywt_kg <- c(
  1.4, 465, 36.3, 27.7, 1., 11700, 2547, 187.1,
  521, 10, 3.3, 529, 207, 62, 6654, 9400,
  6.8, 35, 0.1, 0.02, 2.5, 55.5, 100, 52.2,
  0.3, 87000, 0.1, 192
)

brainwt_kg <- c(
  0.0081, 0.423, 0.1195, 0.115, 0.0055, 0.05, 
  4.603, 0.419, 0.655, 0.115, 0.0256, 0.68, 
  0.406, 1.32, 5.712, 0.07, 0.179, 0.056,
  0.001, 0.0004, 0.0121, 0.175, 0.157, 0.44,
  0.0019, 0.1545, 0.003, 0.18
)
```

## Working with the vectors

- **Check which of the following animals are contained in the `species` vector**

```{r echo=TRUE}
animals_to_check <- c("Snail", "Goat", "Chimpanzee", "Rat", "Dragon", "Eagle")

# Goes through every element in animals_to_check and returns TRUE
# if it appears in species
animals_to_check %in% species
```

You can see that Goat, Chimpanzee and Rat are contained in `species`.

:::{.callout-note}

If you use `%in%` to check if elements are contained in a vector, the result vector is of the same length as the vector that is **before** the `%in%`. 

Compare the result from above with the following:

```{r}
# Goes through every element in species and returns TRUE
# if it appears in animals_to_check
species %in% animals_to_check
```

:::

- **Calculate mean and standard deviation of brain weight**

```{r echo=TRUE}
mean(brainwt_kg)
sd(brainwt_kg)
```

- Which species have a brain weight larger than the mean brain weight of all species?

```{r echo=TRUE}
species[brainwt_kg > mean(brainwt_kg)]
```

- **Calculate the ratio brain weight / body weight in percent and save it in a vector**

```{r echo=TRUE}
brain_body_ratio <- brainwt_kg / bodywt_kg * 100
```

- **Are there any animals with a larger brain to body weight ratio than humans? If yes, which ones?**

**Step 1:** Look at the brain to body ratio of humans

```{r echo=TRUE}
brain_body_ratio[species == "Human"]
```

**Step 2**: Compare it with the other animals

```{r echo=TRUE}
# New variable for human brain to body ratio
bbr_human <- brain_body_ratio[species == "Human"]
# Are there animals that have a larger brain to body ratio than humans?
brain_body_ratio > bbr_human
# Which are these animals
species[brain_body_ratio > bbr_human]

# or short
species[brain_body_ratio > brain_body_ratio[species == "Human"]]
```

## Extras

- **Round the vector to 4 decimal places:**

```{r echo=TRUE}
brain_body_ratio <- round(brain_body_ratio, digits = 4)
```


- **Which animal has the smallest brain to body ratio?**

```{r echo=TRUE}
species[brain_body_ratio == min(brain_body_ratio)]
```

The `r species[brain_body_ratio == min(brain_body_ratio)]` has a really small brain with a weight of just `r brainwt_kg[species == species[brain_body_ratio == min(brain_body_ratio)]]` kg at a body weight of `r bodywt_kg[species == species[brain_body_ratio == min(brain_body_ratio)]]` kg!

- **Add elements to the vector**

```{r echo=TRUE}
species_new <- c("Eagle", "Snail", "Lion")
brainwt_kg_new <- c(0.0004, NA, 0.5)
bodywt_kg_new <- c(18, 0.01, 550)

species <- c(species, species_new)
brainwt_kg <- c(brainwt_kg, brainwt_kg_new)
bodywt_kg <- c(bodywt_kg, bodywt_kg_new)
```

- **What is the mean brain weight now?**

```{r echo=TRUE}
mean(brainwt_kg)
# na.rm = TRUE removes missing values before calculating the mean
mean(brainwt_kg, na.rm = TRUE)
```
