---
title: "Solution: Import and Export Data"
code-line-numbers: false
---

```{r}
#| include: false
library(tidyverse)
library(readxl)
```

## Read a CSV file

```{r}
#| eval: false
library(tidyverse)
```

Read the tree data:

```{r}
#| message: false
trees <- read_csv("data/tree_growth.csv")
trees
```

Explore the data:

```{r}
summary(trees)
```

There are 4 tree species (Beech, Birch, Oak, Pine). The tallest tree is 20.1 m.

Calculate the mean height:

```{r}
mean(trees$height_m)
```

## Read an Excel file

```{r}
#| eval: false
library(readxl)
```

```{r}
birds <- read_excel("data/bird_observations.xlsx")
birds
```

```{r}
summary(birds)
```

## Write data to a file

```{r}
#| eval: false
write_csv(trees, file = "data/trees_copy.csv")
```

:::{.callout-note}

For this to work, the `data/` directory must already exist in your project. Otherwise you will get an error.

:::

## Challenge: a slightly messy file

If you try to read the file directly, the metadata lines on top are treated as data:

```{r}
#| eval: false
read_csv("data/water_quality.csv")
```

Open the file in a text editor — you can see 4 lines of metadata before the actual data starts. Use `skip` to skip them:

```{r}
water <- read_csv("data/water_quality.csv", skip = 4)
water
```

## Optional tasks

### Challenge: an even messier file

This file has two problems:

1. Metadata lines on top (4 lines to skip)
2. The delimiter is `;` instead of `,`

So `read_csv()` won't work even with `skip`. Use `read_delim()` instead:

```{r}
soil <- read_delim(
  "data/soil_nutrients_messy.csv",
  delim = ";",
  skip = 4
)
soil
```

### Clean messy column headers

Some column names have spaces and special characters. Use `janitor::clean_names()` to fix this:

```{r}
#| eval: false
# install.packages("janitor")
library(janitor)
soil <- clean_names(soil)
soil
```

The column names are now clean without spaces and special characters, all lowercase.
