---
title: "Solution to statistical test task"
code-line-numbers: false
---

## Compare the flipper length of penguins using statistical tests

First, create subsets of the flipper length of all species as vectors:

```{r echo=TRUE}
#| message: false
library(palmerpenguins)
library(tidyverse)
adelie <- filter(penguins, species == "Adelie")$flipper_length_mm
gentoo <- filter(penguins, species == "Gentoo")$flipper_length_mm
chinstrap <- filter(penguins, species == "Chinstrap")$flipper_length_mm
```

### Step 1: Test all groups for normality

**With Shapiro-Wilk**

The Shapiro-Wilk test suggests that Adelie and Chinstrap are normally distributed, but Gentoo is not.

```{r, echo=TRUE}
shapiro.test(adelie)
shapiro.test(gentoo)
shapiro.test(chinstrap)
```

**With QQ-plots**

```{r, echo=TRUE}
#| warning: false
library(patchwork) # using patchwork to combine multiple plots into one
ch <- ggpubr::ggqqplot(chinstrap) + labs(title = "chinstrap")
ad <- ggpubr::ggqqplot(adelie) + labs(title = "adelie")
ge <- ggpubr::ggqqplot(gentoo) + labs(title = "gentoo")

ch + ge + ad
```

### Compare Chinstrap vs. Adelie

Both are normally distributed, so we can compare variances for Chinstrap and Adelie penguins:

```{r, echo=TRUE}
var.test(chinstrap, adelie)
```

This indicates that the variances do not differ between Chinstrap and Adelie penguins.
Therefore, we can use a t-test to compare mean flipper length:

```{r echo=TRUE}
t.test(chinstrap, adelie)
```

The t-test suggests, that the mean flipper length differs between Chinstrap and Adelie penguins.

### Compare Chinstrap vs. Gentoo

Gentoo is not normally distributed, so we can use a Wilcoxon-rank-sum test to compare
means:

```{r, echo=TRUE}
wilcox.test(chinstrap, gentoo)
```

The Wilcoxon tests suggests that the flipper lengths differ between Chinstrap and Gentoo penguins.

### Compare Adelie vs. Gentoo

Gentoo is not normally distributed, so we can use a Wilcoxon-rank-sum test to compare
means:


```{r, echo=TRUE}
wilcox.test(adelie, gentoo)
```

The Wilcoxon tests suggests that the flipper lengths differ between Adelie and Gentoo penguins.


### Visual comparison

The boxplot supports the findings of the statistical tests. Notches do not overlap.

```{r, echo=TRUE}
#| warning: false
penguins |> 
  ggplot(aes(x = species, y = flipper_length_mm)) +
  geom_boxplot(notch = TRUE)
```

Similarly the standard errors of the means do not overlap between the species.

```{r echo = TRUE}
#| warning: false
penguins |>
  ggplot(aes(x = species, y = flipper_length_mm)) +
  stat_summary() 
```

To add the significance levels of the tests performed, I used
the `geom_signif()` layer provided by the `ggsignif` package.

Since I performed two different tests (t-test for the comparison of Chinstrap and
Adelie penguins, Wilcoxon test for the other two comparisons), I added two 
different significance layers that distinguish the two tests.

For the boxplot:

```{r echo = TRUE}
#| warning: false
library(ggsignif)
penguins |> 
  ggplot(aes(x = species, y = flipper_length_mm)) +
  geom_boxplot(notch = TRUE) +
  geom_signif(
    comparisons = list(
      c("Chinstrap", "Adelie")
    ),
    test = "t.test",
    test.args = list(var.equal = TRUE),
    map_signif_level = TRUE
  ) +
  geom_signif(
    comparisons = list(
      c("Chinstrap", "Gentoo"),
      c("Gentoo", "Adelie")
    ),
    test = "wilcox.test",
    y_position = c(235, 240),
    map_signif_level = TRUE
  )
```

For a plot with mean and errorbar:

```{r echo =TRUE}
#| warning: false
penguins |> 
  ggplot(aes(x = species, y = flipper_length_mm)) +
  stat_summary() +
  geom_signif(
    comparisons = list(
      c("Chinstrap", "Adelie")
    ),
    test = "t.test",
    test.args = list(var.equal = TRUE),
    map_signif_level = TRUE,
    y_position = 200,
    tip_length = 0.01
  ) +
  geom_signif(
    comparisons = list(
      c("Chinstrap", "Gentoo"),
      c("Gentoo", "Adelie")
    ),
    test = "wilcox.test",
    y_position = c(216, 218),
    map_signif_level = TRUE,
    tip_length = 0.01
  )
```

