### 1.1 and 1.2 Getting started with ggplot and palmerpenguins

```{r eval=TRUE, echo=TRUE}
# install.packages("tidyverse")
# install.packages("palmerpenguins")
library(tidyverse)
library(palmerpenguins)
```
- Have a look at the penguin data set

```{r echo=TRUE, eval=FALSE}
penguins
```
```{r echo=FALSE, eval=TRUE}
theme_set(theme_gray())
print(penguins, n=6)
```


### 1.3 Exploratory plotting

#### 1.3.1 Relationship between bill length and bill depth (scatterplot)

What is the relationship between bill length and bill depth?

- Create a scatterplot with bill length on the x-axis and bill depth on the y-axis
- Can you add a regression line?
- Add species as color aesthetic. Does your interpretation of the data change?
  - What is the difference between adding color as a global aesthetic or as a local aesthetic of the point layer?
- Add species as shape aesthetic as well to distinguish the species

```{r echo=TRUE, eval=TRUE}
# Bill length vs. bill depth scatterplot with regression line
ggplot(
  data = penguins,
  aes(
    x = bill_length_mm,
    y = bill_depth_mm
  )
) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
```{r echo = TRUE, eval=FALSE}
# or short
ggplot(penguins, aes(bill_length_mm, bill_depth_mm)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

If we add the color aesthetic locally to the point layer, only this layer will be affected by it. The regression line is not separated by species but still calculated and plotted for all data points together:

```{r echo=TRUE, eval=TRUE}
# Bill length vs. bill depth scatterplot with regression line
# color as aesthetic local to the point layer
ggplot(penguins, aes(bill_length_mm, bill_depth_mm)) +
  geom_point(aes(color = species)) +
  geom_smooth(method = "lm", se = FALSE)
```

We can see an example of the Simpson's paradox here. If you don't consider species, it looks like the bill depth decreases with bill length. But after separating the data by species, we see that the effect is actually the opposite.

To draw separate regression lines for the species, we need to either add the color aesthetic to the smooth layer as well, or define the color aesthetic gloablly in the top layer `ggplot` call.


```{r echo = TRUE}
# Option A: Define color aesthetic in smooth layer as well
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(color = species)) +
  geom_smooth(aes(color = species),
    method = "lm", se = FALSE
  )
```
```{r eval=FALSE}
# Option B: Define color aesthetic once globally
ggplot(penguins, aes(
  x = bill_length_mm,
  y = bill_depth_mm,
  color = species
)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

#### 1.3.2 Difference in flipper length between species (boxplot)

Is there a difference in flipper length between the species?

- Create a boxplot of the flipper length (y-axis) for the different species (x-axis)
- Try adding notches to the boxplots by setting `notch = TRUE` in the boxplot layer
- **Extra:** Add a layer with `geom_point()`. Try setting `position = position_jitter()` as argument in the point layer. What does it do?

```{r echo = TRUE}
# Basic boxplot of flipper length with notches
ggplot(penguins, aes(species, flipper_length_mm)) +
  geom_boxplot(notch = TRUE)
```
A `geom_point` with `position = position_jitter()` will add the individual data points to the plot. It's important to set a seed here to get the same result for the point position on the x-axis every time. Otherwise your plot is not reproducible.

```{r echo = TRUE}
ggplot(penguins, aes(species, flipper_length_mm)) +
  geom_boxplot() +
  geom_point(position = position_jitter(seed = 123))
```


#### 1.3.3 Differences between body mass of male and female penguins (boxplot)

Are male penguins heavier than female penguins? And is this different between the 3 species?

- Create a boxplot with body mass on the y axis and sex on the x axis
- Add the difference between penguin species to this plot. Try the different options ggplot offers
  - species as color aesthetic
  - species as fill aesthetic
  - species as facet using `facet_wrap`
- **Extra:** What happens if you use `geom_violin` instead of `geom_boxplot`? Can you combine both geoms in one plot?

```{r echo = TRUE}
# Basic boxplot of body mass for penguins of different sex
ggplot(penguins, aes(x = sex, y = body_mass_g)) +
  geom_boxplot()
```
Species as color aesthetic:

```{r echo = TRUE}
ggplot(penguins, aes(x = sex, 
                     y = body_mass_g)) +
  geom_boxplot(aes(color = species))
```
Species as fill aesthetic

```{r echo = TRUE}
ggplot(penguins, aes(x = sex, y = body_mass_g)) +
  geom_boxplot(aes(fill = species))
```

Species as facets:

```{r echo = TRUE}
ggplot(penguins, aes(x = sex, y = body_mass_g)) +
  geom_boxplot() +
  facet_wrap(~species)
```
With geom_violin

```{r echo = TRUE}
ggplot(penguins, aes(x = sex, y = body_mass_g)) +
  geom_violin() +
  geom_boxplot(width = .4) +
  facet_wrap(~species)
```


### 1.4 Beautify the plots

#### 1.4.1 Beatuify plots from 1.3

Here are just some examples of how to make the plots from before prettier. Of course there a many other options as well.

**Example one: Boxplot of flipper length and species**

```{r echo = TRUE}
ggplot(penguins, aes(species, flipper_length_mm, color = species)) +
  geom_boxplot(width = 0.3) +
  geom_point(
    alpha = 0.5,
    position = position_jitter(width = 0.2, seed = 123)
  ) +
  ggsci::scale_color_uchicago() +
  labs(x = "Species", y = "Flipper length (mm)") +
  theme_minimal() +
  theme(legend.position = "none")
```

What was changed compared to the basic plot?

- Add color for each species by setting a global `color` aesthetic
- Make boxes and jitter points less wide by setting `width` for both layers
- Make jitter points slightly transparent by specifying `alpha = 0.5` for the jitter layer
- Change the color to nicer colors from the `ggsci` package
- Change from default theme to `theme_minimal()`
- Remove the legend with `legend.position = "none"`
- Change the axis labels with `labs()`


**Example two: Reproducing the plot from the presentation**

The following code is adapted from the [palmerpengins package website](https://allisonhorst.github.io/palmerpenguins/articles/examples.html).

```{r eval=TRUE, echo=TRUE}
ggplot(
  data = penguins,
  aes(
    x = bill_length_mm,
    y = bill_depth_mm,
    color = species,
    shape = species
  )
) +
  geom_point(size = 3, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_manual(values = c("darkorange", "purple", "cyan4")) +
  labs(
    title = "Penguin bill dimensions",
    subtitle = "Bill length and depth for Adelie, Chinstrap and Gentoo Penguins at Palmer Station LTER",
    x = "Bill length (mm)",
    y = "Bill depth (mm)",
    color = "Penguin species",
    shape = "Penguin species"
  ) +
  theme_minimal() +
  theme(
    legend.position = c(0.85, 0.15),
    legend.background = element_rect(fill = "white", color = NA)
  )
```

What was changed compared to the basic plot?

- Make points larger and slightly transparent by setting `size` and `alpha` for the point layer
- Change to custom color scale
- Add title and subtitle with `labs`
- Change title of x-axis, y-axis and legend for color and shape aesthetic with `labs`
- Use `theme_minimal()` instead of default theme
- Change legend position to bottom right corner within the plot
  - Positions are relative to the bottom left corner of the plot
  - 0.85 (85% of plot width) to the right
  - 0.15 (15% of plot height) towards the top

### 1.5 Save one of the plots on your machine

Example with one of the plots from above:

```{r echo = TRUE}
flipper_box <- ggplot(penguins, aes(species, flipper_length_mm, color = species)) +
  geom_boxplot(width = 0.3) +
  geom_jitter(alpha = 0.5, position = position_jitter(width = 0.2, seed = 123)) +
  ggsci::scale_color_uchicago() +
  labs(x = "Species", y = "Flipper length (mm)") +
  theme_minimal() +
  theme(legend.position = "none")
```
```{echo= TRUE, eval=FALSE}
# save as png in /img directory of the project
ggsave(filename = "./img/flipper_box.png", flipper_box)
# save as pdf in /img directory of the project
ggsave(filename = "./img/flipper_box.pdf", flipper_box)
```


### Some more examples

**Histogram**

```{r echo = TRUE}
ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(alpha = 0.6) +
  ggsci::scale_fill_d3() +
  labs(
    y = "Frequency",
    x = "Flipper length [mm]",
    fill = "Penguin species"
  ) +
  theme_minimal()
```

**Heat map**

For the penguin data set a heat map does not make that much sense. But an example would be:

```{r echo=TRUE}
ggplot(penguins, aes(
  x = species,
  y = sex,
  fill = flipper_length_mm
)) +
  geom_tile() +
  scale_fill_viridis_c() +
  theme_classic()
```

### The `plotly` package

You can use the `plotly package` to create interactive plots where you can e.g. hover over an area of your plot with the mouse and the respective values will be displayed.

To get started, you have to install the plotly package

```{r echo=TRUE, eval=FALSE}
install.packages("plotly")
```

You can use any ggplot and wrap it in the `ggplotly` function from the `plotly` package to make it interactive, e.g.

```{r echo=TRUE}
plotly::ggplotly(flipper_box)
```

### References

Check out the [package website](https://allisonhorst.github.io/palmerpenguins/articles/examples.html) of the `palmerpenguin` package. They have more nice examples of data visualizations that you can do with ggplot.

Horst AM, Hill AP, Gorman KB (2020). palmerpenguins: Palmer Archipelago (Antarctica) penguin data. R package version 0.1.0. https://allisonhorst.github.io/palmerpenguins/. doi: 10.5281/zenodo.3960218.

