```{r}
library(tidyverse)
library(palmerpenguins)
```

#### Filter

Find all penguins that 

- live on the island Dream in the year 2007

```{r}
filter(penguins, island == "Dream" & year == 2007)
```
- have a bill length between 40 and 45 mm

```{r}
filter(penguins, between(bill_length_mm, 40, 45))
# same as 
filter(penguins, bill_length_mm < 45 & bill_length_mm > 40)
```
- for which we know the sex

```{r}
filter(penguins, !is.na(sex))
```
- of the species Adelie or Gentoo

```{r}
filter(penguins, species %in% c("Adelie", "Gentoo"))
# or
filter(penguins, species == "Adelie" | species == "Gentoo")
```

#### count

- count the penguins on each island

```{r}
count(penguins, island)
```
- count the penguins on each island and of each species

```{r}
count(penguins, island, species)
```

#### Arrange

- sort `penguins` to find penguins with the highest body mass

```{r}
arrange(penguins, desc(body_mass_g))
```

- sort `penguins` to find penguins with the lowest body mass
```{r}
arrange(penguins, body_mass_g)
```

- sort `penguins` by species and sex

```{r}
arrange(penguins, species, sex)
```

#### Select

- select only the columns species, sex and year

```{r}
select(penguins, species, sex, year)
```

- select the variables based on the following vector

```{r error=TRUE}
cols <- c("species", "bill_length_mm", "flipper_length_mm", "i_dont_exits")
select(penguins, any_of(cols))
# this would return  an error
select(penguins, all_of(cols))
```
- select only columns that contain measurements in mm

# select
pick columns based on their name

helpers
starts_with()
ends_with()
contains()
matches()
from:to

# arrange

desc()

# mutate
multiple columns
ifelse
case_when

# transmute

# summarize

group_by
ungroup

# pipe
