## Getting started with tibbles

Before you start the task:

- Install the tibble package by calling `install.packages("tibble")` in the console.
- Open your script for the task and put `library(tibble)` on the top
- Execute this line and the tibble package is loaded into your R session and ready to use

## Creating a tibble

Use the vectors `species`, `brainwt_kg`, `bodywt_kg` from the last task and put them in a tibble.

Add several columns to this tibble:

- The ratio of brain weight to body weight
- A column that indicates whether an animal has a large or a small brain
  - Put `"large"` into the column if the brain-body weight ratio is higher than 1%
  - Else put `"small"` into the column
- Add the order of the animals to the data:

```{r eval=FALSE}
order <- c(
  "Rodentia", "Artiodactyla", "Carnivora", "Artiodactyla", "Rodentia", NA,
  "Proboscidea", rep("Perissodactyla", 2), NA, "Carnivora", "Artiodactyla",
  rep("Primates", 2), "Proboscidea", "Ornithischia", "Primates", "Diprotodontia",
  "Rodentia", "Rodentia", "Lagomorpha", "Artiodactyla", "Carnivora", "Primates",
  "Rodentia", NA, "Eulipotyphla", "Artiodactyla"
)
```



## Functions

```{r}
install.packages()
library()
tibble()
as_tibble()
str()
nrow()
ncol()
names()
tibble::view()
summary()
```


## Indexing

```{r}
tibble[row, column]
tibble[columns]
# by column name or column position
$
# logical indexing with logical, relational operators, %in%
```

## Wroking with tibbles

- replace values
Adding new columns:
- ifelse
- columns based on other columns

# summary

- data frames and tibbles are named lists of vectors of the same length
- basic props: name (header), same data type, same length
- tibbles are better than df but you can do same wiht them
