## Creating a tibble

Use the vectors `species`, `brainwt_kg`, `bodywt_kg` from the last task and put them in a tibble.

```{r}
# install.packages("tibble")
library(tibble)

species <- c(
  "MountainBeaver", "Cow", "GreyWolf", "Goat",
  "GuineaPig", "Diplodocus", "AsianElephant", "Donkey",
  "Horse", "PotarMonkey", "Cat", "Giraffe",
  "Gorilla", "Human", "AfricanElephant", "Triceratops",
  "RhesusMonkey", "Kangaroo", "GoldenHamster", "Mouse",
  "Rabbit", "Sheep", "Jaguar", "Chimpanzee",
  "Rat", "Brachiosaurus", "Mole", "Pig"
)

bodywt_kg <- c(
  1.4, 465, 36.3, 27.7, 1., 11700, 2547, 187.1,
  521, 10, 3.3, 529, 207, 62, 6654, 9400,
  6.8, 35, 0.1, 0.02, 2.5, 55.5, 100, 52.2,
  0.3, 87000, 0.1, 192
)

brainwt_g <- c(
  8.1, 423, 119.5, 115, 5.5, 50, 4603, 419, 655, 115, 25.6,
  680, 406, 1320, 5712, 70, 179, 56, 1, 0.4, 12.1, 175,
  157, 440, 1.9, 154.5, 3, 180
)
# Creating the tibble 
animals <- tibble(
  species = species,
  bodywt_kg = bodywt_kg,
  brainwt_kg = brainwt_g / 1000
)
```


Adding several columns to this tibble:


```{r}
# Add brain to body ratio (%)
animals$bbr <- animals$brainwt_kg / animals$bodywt_kg * 100
# Add the brain size category
animals$brainwt_category <- ifelse(
  animals$bbr > 1,
  "large",
  "small"
)
# Add the order of the species
order <- c(
  "Rodentia", "Artiodactyla", "Carnivora", "Artiodactyla", "Rodentia", NA,
  "Proboscidea", rep("Perissodactyla", 2), NA, "Carnivora", "Artiodactyla",
  rep("Primates", 2), "Proboscidea", "Ornithischia", "Primates", "Diprotodontia",
  "Rodentia", "Rodentia", "Lagomorpha", "Artiodactyla", "Carnivora", "Primates",
  "Rodentia", NA, "Eulipotyphla", "Artiodactyla"
)
animals$order <- order
```

## Exploring the tibble

```{r}
view(animals)
summary(animals)
str(animals)
nrow(animals)
ncol(animals)
names(animals)
```

## Indexing

Warm up

```{r}
animals[c(1,5,7), c("species", "bodywt_kg")]
animals[1:10, ]
# last row last column without using the index
animals[nrow(animals), ncol(animals)]
# select specific orders
orders_select <- c("Primates", "Rodentia", "Carnivora")
animals[animals$order %in% orders_select,]
```
Create two subsets for large and small brained animals and calculate the mean and standard deviation of the brain to body ratio in these subsets.

Create subsets

```{r}
large_brain <- animals[animals$brainwt_category == "large", ]
small_brain <- animals[animals$brainwt_category == "small", ]
```


Calculate the means and standard deviation

```{r}
# Large brains
mean(animals[animals$brainwt_category == "large", ]$bbr) # or

mean(large_brain$bbr)
sd(large_brain$bbr)

# small brain
mean(animals[animals$brainwt_category == "small", ]$bbr) # or

mean(small_brain$bbr)
sd(small_brain$bbr)
```

