---
title: "Gapminder Analysis"
author: "Selina Baldauf"
date: today
format:
  html:
    toc: true
    toc-location: left
    number-sections: true
    code-fold: true
execute:
  warning: false
  message: false
---

```{r}
#| include: false
library(gapminder)
library(ggplot2)
library(dplyr)
```

## Introduction

The **gapminder** dataset contains data about *life expectancy*, *GDP per capita*, and *population* for countries around the world.

In this analysis, we will:

- Load the gapminder data
- Look at life expectancy in 2007
- Create a plot of life expectancy vs. GDP per capita

You can find more information about the data on the [Gapminder website](https://www.gapminder.org/).

## Analysis

The gapminder data contains `r nrow(gapminder)` observations from `r length(unique(gapminder$country))` countries.

```{r}
#| label: fig-life-exp
#| fig-cap: "Life expectancy vs. GDP per capita in 2007"
#| fig-align: center
#| out-width: "80%"

gapminder_2007 <- gapminder |>
  filter(year == 2007)

ggplot(gapminder_2007, aes(x = gdpPercap, y = lifeExp)) +
  geom_point(aes(color = continent)) +
  theme_minimal()
```

As we can see in @fig-life-exp, life expectancy generally increases with GDP per capita.

### Summary by continent

```{r}
gapminder_2007 |>
  group_by(continent) |>
  summarize(
    n_countries = n(),
    mean_life_exp = round(mean(lifeExp), 1),
    mean_gdp = round(mean(gdpPercap), 0)
  ) |>
  knitr::kable()
```

### Statistical test

Is the mean life expectancy different between Europe and Africa?

```{r}
europe <- gapminder_2007 |>
  filter(continent == "Europe") |>
  pull(lifeExp)

africa <- gapminder_2007 |>
  filter(continent == "Africa") |>
  pull(lifeExp)

t.test(europe, africa) |>
  broom::tidy() |>
  knitr::kable()
```
